### 2022-02-25 introduction.tex
Some introduction/background added.

\chapter{Introduction\label{cha:introduction}}

Solar power production is extremely dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016}. A single cloud can seriously impact the production of a solar farm. Such unreliability in a power production system is highly undesirable, since production must match demand \cite{lee_forecasting_2018}. When solar power production falls, the power demand it was serving must be fulfilled by other means. Fossil fueled power plants are generally kept on standby, waiting to fulfill this demand \cite{lee_forecasting_2018}. These plants often take hours to start, generally pollute more than other plants and are very expensive to run \cite{lee_forecasting_2018}.
Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}. This notice enables planning for the dips in power production and making arrangements into the future. This provides the tools to reduce the cost and pollution of power generation as well as increasing the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
Recent advancements in deep learning have unlocked new tools that have been shown to be extremely effective in predicting the future behavior of a system given enough relevant data.\\

%% \ifdraft only shows the text in the first argument if you are in draft mode.
%% These directions will disappear in other modes.
\ifdraft{State the objectives of the exercise. Ask yourself:
  \underline{Why} did I design/create the item? What did I aim to
  achieve? What is the problem I am trying to solve?  How is my
  solution interesting or novel?}{}

\section{Background}

Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time based data efficiently. LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.\\
Lin et al. \cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on the task, Temporal Convolutional Neural Networks (TCNN) can show even better performance. \textit{"TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
and residual connections"} \cite{lin_temporal_2020}. \\
Jaidee et al. \cite{jaidee_very_2019} showed that for very short term predictions, on the time scale of a few hours, LSTMs, GRUs and derivative methods show relatively similar performance. This timescale is where it is critical to accurately predict cloud movements. 
These methods have been taken as far as building genetic algorithms to generate the optimal neural network for the purpose \cite{jaidee_very_2019}. These genetic algorithms operate on the same principals as animal breeding. They generate multiple candidate neural networks and use the best networks as a base to generate new networks for the next generation. This is done for multiple generations and in the end the best neural network generated is used \cite{jaidee_very_2019}.\\
Su et al. \cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto Regressive Neural Network (NARXNN), as well as various statistical approaches. When predicting power output, NARXNN takes in the current power output values and considers them as well as the past values of the power output of the system. These experiments showed that NARXNN had the best performance of these, by a rather large margin, and a hybrid method, of the better performing methods, showed even better results \cite{anderson_using_2018}. The statistical methods generally performed significantly worse in predicting the power output. \\
Satellite imagery has been used for predicting solar power output as well. A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.\\

\ifdraft{Provide background about the subject matter (e.g. How was morse code
developed?  How is it used today?). 
This is a place where there are usually many citations.
It is suspicious when there is not.
Include the purpose of the different equipment and your design intent. 
Include references to relevant scientific/technical work and books.
What other examples of similar designs exist?
How is your approach distinctive?

If you have specifications or related standards, these must be
described and cited also.  As an example, you might cite the specific
RoboSub competition website (and documents) if working on the lighting system for an AUV\cite{guls2016auvlight}

\section{Transformers}
Hard to train
sequential
 -> parallelization hard
encoder/decoder instead. Attention. 
 Explain enc/dec
 Autoencoders generate same output as the input, so no labeled data is needed -> completely unsupervisd learning.
 Locality a problem with CNNs and RNNs. Can easily find local context but keeping on to context far apart in input is difficult. Enter transformers.
 Encoder reads blocks of data and uses that and previously generated data as inputs into decoder which generatesdata one point at a time.
 Attention. How you relate things to each other. i.e. input to output
 self attention. How do features in the input data relate to each other

\subsection{Encoder/decoder}

\subsection{Attention}

\section{Goals}

%% Glossary is broken, do not use --foley
% \gls{auv}\footnote{Autonomous Undersea Vehicle}.

% Notice that there is now information on the AUV in the Index and Acronyms.
% It isn't in the \gls{glossary} because we didn't put it there.
\index{AUV}
}{}

### 2022-04-05 code.tex
Results and conclusion

\chapter{Code}\label{cha:code}
You can put code in your document using the listings package, which is
loaded by default in \path{custom.tex}.  Be aware that the listings
package does not put code in your document if you are in draft mode
unless you set the \texttt{forcegraphics} option.

There is an example java (Listing~\ref{src:Data_Bus.java}) and XML
file (Listing~\ref{src:AndroidManifest.xml}).  Thanks to the
\texttt{url} package, you can typeset OSX and unix paths like this:
\path{/afs/rnd.ru.is/project/thesis-template}.  Windows paths:
\path{C:\windows\temp\ }.  You can also typeset them using the menukey
package, but it tends to delete the last separator and has other
complications.\footnote{The menukey package has issues with biblatex,
  read \path{custom.tex} for more information.}

If you are trying to include multiple different languages, you should
go read the documentation and set these up in \path{custom.tex}.  You
will save yourself a lot of effort, especially if you have to fix
anything.

%I have put the source code in the \directory{src/} folder.
%\lstinputlisting[language=Java, firstline=1,
%lastline=40, caption={Data\_Bus.java: Setting up the class.},
%label={src:Data_Bus.java}]{src/Data_Bus.java}

%\lstinputlisting[language={[android]XML}, firstline=1, lastline=20,
%caption={AndroidManifest.xml: Configuration for the Android UI.},
%label={src:AndroidManifest.xml}]{src/AndroidManifest.xml}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "DEGREE-NAME-YEAR"
%%% End:

### 2022-04-05 discussion.tex
Results and conclusion

\chapter{Discussion}
%In this section you discuss any issues that came up while developing
%the system.  If you found something particularly interesting,
%difficult, or an important learning experience, put it here.  This is
%also a good place to put additional figures and data.
Mean Absolute Percentage Error (MAPE) does not tell the whole story of a model's performance. The true measure of performance is how useful the model output is to energy producers. The model's confidence in a prediction is almost as important as the accuracy of the prediction and the total energy produced over the day matters.
The forecast downward shortwave irradiance, the largest factor in measured irradiance, rarely drops. It is almost exclusively on very high variance days that it deviates from a sine wave. The downward diffuse shortwave irradiance, the other network input data value, is a much smaller part of the total irradiance and has, to the bare eye, seemingly no correlation with the observation values. Despite this, the models show generally fantastic performance in predicting the irradiance.
The full model was able to utilize the forecast data from the surrounding area to achieve noticeably better performance. Of course, since the limited model uses only one ninth the data of the full model, some hyperparameter tuning had to be done to optimize its performance and some of the performance difference could be explained by worse optimization.
With the data at hand, it is not possible to separate the performance of the weather forecast model producing our input data and the performance of our models. We do not know how close this model comes to maximizing the information available in the input data.

Fig.~\ref{fig:disc_low} shows side-by-side the two models' performance for the same calm day. The full model was able to track the observations more closely with more confidence. This was typical for all calm days.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/disc/disc_med_full.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/low/low_g_1.png}}\qquad
    \caption{Comparison of model performance on a clear day.
    \label{fig:disc_low}}
\end{figure}

In Fig.~\ref{fig:disc_med} we can see a comparison between the two models on a less predictable day. Again, the full model shows better performance in tracking, primarily because the limited model misses an end-of-day prediction. There are similar days where the limited model shows slightly better performance.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_3.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}}\qquad
    \caption{Comparison of model performance on a day with some drops in irradiance.
    \label{fig:disc_med}}
\end{figure}

Fig.~\ref{fig:disc_high} shows a clear example of the performance difference between the models on erratic days. The full model, while still having a high error gets somewhat close to keeping the observations within its much smaller 90\% confidence interval than the limited model. The limited model also predicts much more irradiance over the day than the, already too high, full model.
\begin{figure}[ht!]
    \centering
    \subfloat[Full model]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_3.png}}\qquad
    \subfloat[Limited model]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_4.png}}\qquad
    \caption{Comparison of model performance on an erratic day.
    \label{fig:disc_high}}
\end{figure}

\section{Conclusion\label{sec:conclusions}}
The results clearly show that relying forecast irradiance data from meteorological models is a very effective, but not perfect, method of accurately predicting irradiance in a single area. A transformer model, the Temporal Fusion Transformer, managed to very effectively translate patterns in meteorological irradiance forecasts to accurate local irradiance predictions utilizing spatially and temporally adjacent data.

We used a dataset of a fairly limited size, only 425 days. A few years of data for training the model would likely yield somewhat better results. The data can be processed more before being input into the network to account for known variables like daily and yearly swings in irradiance, which might improve performance. There exists satellite irradiance measurement data which might be used for improving accuracy. It would be interesting to see if a neural network could extract information from the seven pyranometers scattered across the power station, rather than only dropping the highest and lowest values and taking an average.

Sadly, there is not a standard benchmark for solar irradiance forecast models, and the data used by others who have done similar work is not available, so we can't make a proper comparison between methodologies and models. 
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "DEGREE-NAME-YEAR"
%%% End:

### 2022-04-05 introduction.tex
Results and conclusion

@@
 %% These directions will disappear in other modes.
-\ifdraft{State the objectives of the exercise. Ask yourself:
-  \underline{Why} did I design/create the item? What did I aim to
-  achieve? What is the problem I am trying to solve?  How is my
-  solution interesting or novel?}{}
+%\ifdraft{State the objectives of the exercise. Ask yourself:
+%  \underline{Why} did I design/create the item? What did I aim to
+%  achieve? What is the problem I am trying to solve?  How is my
+%  solution interesting or novel?}{}
 
@@
 
+\subsection{LSTMs and GRUs}
 Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
@@
 LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.\\
+
+%\ifdraft{Provide background about the subject matter (e.g. How was morse code
+%developed?  How is it used today?). 
+%This is a place where there are usually many citations.
+%It is suspicious when there is not.
+%Include the purpose of the different equipment and your design intent. 
+%Include references to relevant scientific/technical work and books.
+%What other examples of similar designs exist?
+%How is your approach distinctive?
+
+%If you have specifications or related standards, these must be
+%described and cited also.  As an example, you might cite the specific
+%RoboSub competition website (and documents) if working on the lighting system for an AUV\cite{guls2016auvlight}
+
+\subsection{Transformers}
+%Hard to train
+%sequential
+% -> parallelization hard
+%encoder/decoder instead. Attention. 
+% Explain enc/dec
+% Autoencoders generate same output as the input, so no labeled data is needed -> completely unsupervisd earning.
+% Locality a problem with CNNs and RNNs. Can easily find local context but keeping on to context far apart in input is difficult. Enter transformers.
+ %Encoder reads blocks of data and uses that and previously generated data as inputs into decoder which generatesdata one point at a time.
+ %Attention. How you relate things to each other. i.e. input to output
+ %self attention. How do features in the input data relate to each other
+
+Transformers have been used for a while in Natural Language Processing (NLP) and have more recently been used very effectively in forecasting.\cite{vaswani_attention_2017} They can more effectively learn and preserve local context located far apart in the input data.
+As an example from NLP, the training data can contain multiple cases where time is being told, far apart.
+A transformer model would then more easily identify that the word “o’clock” is generally preceded by a number, whereas RNNs would be more likely to struggle with such distributed information.
+This can be applied to our work as similar weather patterns would reoccur days apart and cause a similar effect on measured irradiance at the power station.
+
+\subsubsection{Encoder/decoder}
+Transformers have an encoder/decoder structure.
+An encoder takes an input sequence and generates an abstract hidden state which the decoder then uses to generate an output sequence.
+A major advantage of this methodology is that the inputs and outputs do not have to have the same length, where the primary limit of the architecture is that all information in the input sequence needs to be represented by a single vector for the decoder to decode.
+The encoder distills the input down to a much smaller representation, a bottleneck, only containing desired information which the decoder in turn extracts.This is on its own used for tasks like de-noising or colorizing images.
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[scale=0.5]{imgs/encoder_decoder.jpeg}
+    \caption{https://towardsdatascience.com/what-is-an-encoder-decoder-model-86b3d57c5e1a
+    \label{fig:encoder_decoder}}
+\end{figure}
+
+\subsubsection{Attention} We previously discussed transformers' ability to preserve context across the long input sequences.
+That is thanks to the attention layer.
+The attention layer encodes an input data point's relevance to another input.
+Calling back to the previous NLP example, it encodes the context that the word "o'clock" is relevant to numbers because in most inputs it is preceded by a number.
+
+Self-attention
+
+Multi-headed attention
+
+\subsubsection{Positional encoding}
+
+\subsection{Weather forecasts}
+
+\subsection{Previous work}
 Lin et al.
@@
 
-\ifdraft{Provide background about the subject matter (e.g. How was morse code
-developed?  How is it used today?). 
-This is a place where there are usually many citations.
-It is suspicious when there is not.
-Include the purpose of the different equipment and your design intent. 
-Include references to relevant scientific/technical work and books.
-What other examples of similar designs exist?
-How is your approach distinctive?
-
-If you have specifications or related standards, these must be
-described and cited also.  As an example, you might cite the specific
-RoboSub competition website (and documents) if working on the lighting system for an AUV\cite{guls2016auvlight}
-
-\section{Transformers}
-Hard to train
-sequential
- -> parallelization hard
-encoder/decoder instead. Attention. 
- Explain enc/dec
- Autoencoders generate same output as the input, so no labeled data is needed -> completely unsupervisd learning.
- Locality a problem with CNNs and RNNs. Can easily find local context but keeping on to context far apart in input is difficult. Enter transformers.
- Encoder reads blocks of data and uses that and previously generated data as inputs into decoder which generatesdata one point at a time.
- Attention. How you relate things to each other. i.e. input to output
- self attention. How do features in the input data relate to each other
-
-\subsection{Encoder/decoder}
-
-\subsection{Attention}
-
 \section{Goals}
+The methods previously used all have in common that they must effectively predict the weather so they can predict power output. 
+Here we explore whether better prediction accuracy can be reached
+if high quality, high resolution weather forecast data is used. This offloads the difficult task of predicting the weather from the responsibilities of the network. \\
+Can we construct a transformer based neural network which, by utilizing weather forecast data, is superior to neural networks which only utilize historical weather data?\\
 
@@
 % It isn't in the \gls{glossary} because we didn't put it there.
-\index{AUV}
-}{}
+%\index{AUV}
+%}{}

### 2022-04-05 methods.tex
Results and conclusion

\chapter{Methods}

\section{Data}
The data used comes from two primary sources
\subsection{Irradiance measurements}
\subsection{Forecast Data}
\subsubsection{Data Source}
Text\cite{gueymard_global_2008, rognvaldsson_numerical_2013, jimenez_wrf-solar_2016}
\subsubsection{Forecast Effectiveness}
\subsection{Error metric}

\section{Tools}

For training models we used a computer with an AMD R9 5950X, 32GB RAM and an Nvidia Geforce RTX 3080Ti. This allowed for quick training of models, as with the relatively small amount of data, training a model took 5-10 minutes.

\section{Model}
    \subsection{Temporal Fusion Transformer}
    Text \cite{lim_temporal_2020}
    \subsection{Setup}
    \subsubsection{Loss function}\label{cha:loss_function}
    Quantile loss and loss quantiles

### 2022-04-05 results.tex
Results and conclusion

\chapter{Results}
Both models showed a strong ability to predict solar irradiance one day into the future.
The full model, using weather forecast data from the surrounding area, showed both a lower Mean Absolute Percentage Error (MAPE), at 13.21\% vs. 15.98\%, as well as better tracking in challenging conditions and a better ability to recognize uncertain conditions, generally reporting lower confidence when its predictions were wrong. Both models can be trained to achieve slightly lower MAPE, but that came at the cost of a higher tendency to underfit to variance, and higher false confidence.
We show irradiance plots for select days representative of the models' performance in various conditions. These plots show the measured irradiance for the day, the models' predicted irradiance values, along with a 90\% confidence interval, detailed in section~\ref{cha:loss_function}. The plots also show the forecast downward shortwave irradiance, the largest factor in measured irradiance, from the model input. Additionally the graphs contain the average MAPE for the day, and the sum of prediction and observation values. The sum values act as a gauge for the total irradiance for the day.

\section{Full model}
The full model, utilizing irradiance data from surrounding areas achieved an average MAPE of 13.21\% on the test dataset. The maximum daily average MAPE was 138.87\% and the minimum was 2.50\%. 
\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.75]{imgs/graphs/full/days_full.png}
    \caption{The recorded Mean Absolute Percentage Error (MAPE) for each day in the test dataset for the full model. 
    \label{fig:days_full}}
\end{figure}

\subsection{Low variance days}
In the test dataset there are a number of clear days, where the sun was mostly or entirely unobstructed. The model predicted most of these days with very low error, in the range 2-5\%. The 90\% confidence interval stayed within $\pm100 W/m^2$ for the whole day and the true prediction was almost spot-on, as can be seen in Fig.~\ref{fig:full_low_best}.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/best_3.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/best_2.png}}\qquad
    \caption{Examples of clear days where the model preformed very well.
    \label{fig:full_low_best}}
\end{figure}

The model did not manage near-perfect predictions on all of these clear days, and made a few under-predictions, as can be seen in Fig.~\ref{fig:full_low_worst}.
These predictions were still even and within the 90\% confidence interval, which was slightly larger than for the other clear day predictions. These predictions had a higher MAPE of 10-13\%.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/worst_1.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/worst_2.png}}\qquad
    \caption{Examples of clear days where the model did not perform well.
    \label{fig:full_low_worst}}
\end{figure}

\subsection{Medium variance days}
On days with some variance in the irradiance measured, the model was generally able to predict the irradiance with relatively good accuracy.
The model has a tendency to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value was within the 90\% confidence interval.

Fig.~\ref{fig:full_med_best} shows examples of good predictions with high confidence, though the confidence is generally lower the farther the irradiance is from the value of a clear day.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_1.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_2.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_3.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_4.png}}\qquad
    \caption{Examples of days with some drops in irradiance, where the model performed very well.
    \label{fig:full_med_best}}
\end{figure}

In Fig.~\ref{fig:full_med_med} we can see examples of days where the model achieves higher error, MAPE in the range 19-35\%, but with generally larger 90\% confidence intervals so the model still keeps the measured values close to or within them.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_m_1.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_m_2.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_m_3.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_m_4.png}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_m_5.png}}\qquad
    \caption{Examples of days with some drops in irradiance, where the model performed somewhat well.
    \label{fig:full_med_med}}
\end{figure}

For a few medium variance days, the model did not track well. Figs.~\ref{fig:full_med_bad_a} and \ref{fig:full_med_bad_d} show where the model predicted a constant lower irradiance when half the day had high irradiance and the other half low. Figs.~\ref{fig:full_med_bad_b} and \ref{fig:full_med_bad_c} show the model predicting a mostly or fully  sunny day when it wasn't. In all of these cases, the observed irradiance was far outside the 90\% confidence interval.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_b_1.png}\label{fig:full_med_bad_a}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_b_2.png}\label{fig:full_med_bad_b}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_b_3.png}\label{fig:full_med_bad_c}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_b_4.png}\label{fig:full_med_bad_d}}\qquad
    \caption{Examples of days with some drops in irradiance, where the model performed badly.
    \label{fig:full_med_bad}}
\end{figure}

\clearpage
\subsection{High variance days}
A few days in the dataset had very low irradiance for part of, or the whole day or large fluctuations in irradiance. These situations were most difficult for the model to predict. Figs.~\ref{fig:full_high_a}, \ref{fig:full_high_b} and \ref{fig:full_high_c} show the model somewhat following the observations, but with a high MAPE of 38-80\% because of offsets in time and/or magnitude. The observed value was often outside the 90\% confidence interval of the predictions. For Figs.~\ref{fig:full_high_a}, \ref{fig:full_high_b} the sum of predicted values is close to the sum of observed values, indicating relatively high accuracy for the days taken as a whole. Fig.\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that receives very little irradiance.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_1.png}\label{fig:full_high_a}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_2.png}\label{fig:full_high_b}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_3.png}\label{fig:full_high_c}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/high/high_4.png}\label{fig:full_high_d}}\qquad
    \caption{Examples of days with large drops in irradiance or low irradiance throughout.
    \label{fig:full_high}}
\end{figure}

\clearpage
\section{Less forecast data}
The limited model, utilizing irradiance data only from the one forecast region, and not surrounding regions, achieved an average MAPE of 15.98\% on the test dataset. The maximum daily average MAPE was 122.87\% and the minimum was 3.47\%. 
\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.75]{imgs/graphs/less/days_single.png}
    \caption{The recorded Mean Absolute Percentage Error for each day in the test dataset of the limited model. 
    \label{fig:days_less}}
\end{figure}

\subsection{Low variance days}
The model predicts most clear days with relatively high accuracy, but even on the best days it deviates from the quite a bit from the observed values, as shown in Fig.~\ref{fig:less_low_a}. The model sometimes under-predicts on clear days. In Fig.~\ref{fig:less_low_b} we can see an example of a day where, even though the 90\% confidence area is large, the observed irradiance is outside it.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/low/low_g_1.png}\label{fig:less_low_a}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/low/low_b_1.png}\label{fig:less_low_b}}\qquad
    \caption{Examples of model performance on clear days.
    \label{fig:less_low}}
\end{figure}

\subsection{Medium variance days}
The model shows less accuracy when the irradiance is not steady. Predictions for most days have the smooth profile of an optimal day, but are scaled to fit closer to the lower total irradiance for the day. The model is inconsistent in how well it tracks the observed irradiance and the confidence it does that with. Sometimes, as shown in Fig.~\ref{fig:less_med_a} the model tracks relatively well, with the observations mostly within the rather tight 90\% confidence interval of roughly $+50/-150 W/m^2$. In Fig.~\ref{fig:less_med_b} we see an example where the model was wrong with very high confidence for the first half of the day. Fig.~\ref{fig:less_med_c} shows the opposite, the model predicts low irradiance for the whole day with a very large 90\% confidence interval of roughly $\pm200 W/m^2$.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}\label{fig:less_med_a}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_2.png}\label{fig:less_med_b}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_1.png}\label{fig:less_med_c}}\qquad
    \caption{Examples of days with some drops in irradiance.
    \label{fig:less_med}}
\end{figure}

\clearpage
\subsection{High variance days}
For the few days in the dataset that had very low irradiance for part of, or the whole day or large fluctuations in irradiance, the model showed high error, MAPE in the range 53-123\% with varying confidence. Figs.~\ref{fig:less_high_a} and \ref{fig:less_high_b} show large over-predictions with high confidence, and Figs.~\ref{fig:less_high_c} and \ref{fig:less_high_d} show days where the 90\% confidence interval covered the top 60\% of the graph, yet it did not cover the observations.
\begin{figure}[ht!]
    \centering
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_1.png}\label{fig:less_high_a}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_2.png}\label{fig:less_high_b}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_3.png}\label{fig:less_high_c}}\qquad
    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/high/high_4.png}\label{fig:less_high_d}}\qquad
    \caption{Examples of days with large drops in irradiance or low irradiance throughout.
    \label{fig:less_high}}
\end{figure}

### 2022-04-08 discussion.tex
A lot of work in methods

@@
 The full model was able to utilize the forecast data from the surrounding area to achieve noticeably better performance.
-Of course, since the limited model uses only one ninth the data of the full model, some hyperparameter tuning had to be done to optimize its performance and some of the performance difference could be explained by worse optimization.
+Of course, since the limited model uses only one-ninth of the data of the full model, some hyperparameter tuning had to be done to optimize its performance and some of the performance difference could be explained by worse optimization.
 With the data at hand, it is not possible to separate the performance of the weather forecast model producing our input data and the performance of our models.
@@
 \section{Conclusion\label{sec:conclusions}}
-The results clearly show that relying forecast irradiance data from meteorological models is a very effective, but not perfect, method of accurately predicting irradiance in a single area.
+The results clearly show that relying on forecast irradiance data from meteorological models is a very effective, but not perfect, method of accurately predicting irradiance in a single area.
 A transformer model, the Temporal Fusion Transformer, managed to very effectively translate patterns in meteorological irradiance forecasts to accurate local irradiance predictions utilizing spatially and temporally adjacent data.
@@
 The data can be processed more before being input into the network to account for known variables like daily and yearly swings in irradiance, which might improve performance.
-There exists satellite irradiance measurement data which might be used for improving accuracy.
+There exists satellite irradiance measurement data that might be used for improving accuracy.
 It would be interesting to see if a neural network could extract information from the seven pyranometers scattered across the power station, rather than only dropping the highest and lowest values and taking an average.

### 2022-04-08 introduction.tex
A lot of work in methods

@@
 A single cloud can seriously impact the production of a solar farm.
-Such unreliability in a power production system is highly undesirable, since production must match demand \cite{lee_forecasting_2018}.
+Such unreliability in a power production system is highly undesirable since production must match demand \cite{lee_forecasting_2018}.
 When solar power production falls, the power demand it was serving must be fulfilled by other means.
-Fossil fueled power plants are generally kept on standby, waiting to fulfill this demand \cite{lee_forecasting_2018}.
+Fossil fueled power plants are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
 These plants often take hours to start, generally pollute more than other plants and are very expensive to run \cite{lee_forecasting_2018}.
 Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}.
-This notice enables planning for the dips in power production and making arrangements into the future.
-This provides the tools to reduce the cost and pollution of power generation as well as increasing the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
-Recent advancements in deep learning have unlocked new tools that have been shown to be extremely effective in predicting the future behavior of a system given enough relevant data.\\
+This notice enables planning for the dips in power production and making arrangements for the future.
+This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
+Recent advancements in deep learning have unlocked new tools that have been shown to be extremely effective in predicting the future behaviour of a system given enough relevant data.\\
 
@@
 \subsection{LSTMs and GRUs}
+\textcolor{blue}{Is there value in going deeper?}\\
 Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
-RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time based data efficiently.
+RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently.
 LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.\\
@@
  %self attention. How do features in the input data relate to each other
+ \textcolor{red}{TODO: Make this coherent}
 
@@
 
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[scale=0.7]{imgs/transformer.png}
+    \caption{The transformer architecture.\cite{vaswani_attention_2017}
+    \label{fig:transformer}}
+\end{figure}
+
 \subsubsection{Encoder/decoder}
@@
 An encoder takes an input sequence and generates an abstract hidden state which the decoder then uses to generate an output sequence.
-A major advantage of this methodology is that the inputs and outputs do not have to have the same length, where the primary limit of the architecture is that all information in the input sequence needs to be represented by a single vector for the decoder to decode.
-The encoder distills the input down to a much smaller representation, a bottleneck, only containing desired information which the decoder in turn extracts.This is on its own used for tasks like de-noising or colorizing images.
+A major advantage of this methodology is that the inputs and outputs do not have to have the same length, whereas the primary limit of the architecture is that all information in the input sequence needs to be represented by a single vector for the decoder to decode.
+The encoder distils the input down to a much smaller representation, a bottleneck, only containing desired information which the decoder in turn extracts.
+This is on its own used for tasks like de-noising or colourizing images.
 
@@
     \includegraphics[scale=0.5]{imgs/encoder_decoder.jpeg}
-    \caption{https://towardsdatascience.com/what-is-an-encoder-decoder-model-86b3d57c5e1a
+    \caption{The encoder/decoder structure
     \label{fig:encoder_decoder}}
@@
 
-\subsubsection{Attention} We previously discussed transformers' ability to preserve context across the long input sequences.
+\subsubsection{Positional encoding}
+Unlike RNNs, the transformer architecture has no recurrence and does inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
+
+\subsubsection{Self-attention} We previously discussed transformers' ability to preserve context across the long input sequences.
 That is thanks to the attention layer.
-The attention layer encodes an input data point's relevance to another input.
-Calling back to the previous NLP example, it encodes the context that the word "o'clock" is relevant to numbers because in most inputs it is preceded by a number.
+The attention layer encodes an input data point's relevance to the rest of the points in the input sequence.
+Calling back to the previous NLP example, it encodes the context that the word "o'clock" is relevant to numbers because it is preceded by a number in most inputs.
 
-Self-attention
-
-Multi-headed attention
-
-\subsubsection{Positional encoding}
+A single attention head can put all the focus on a single element.
+Multiple attention heads can be used to let the transformer model consider the context of more elements at the same time.\cite{rohrer_transformers_2021}
 
 \subsection{Weather forecasts}
+\textcolor{red}{TODO}
 
@@
 Jaidee et al.
-\cite{jaidee_very_2019} showed that for very short term predictions, on the time scale of a few hours, LSTMs, GRUs and derivative methods show relatively similar performance.
+\cite{jaidee_very_2019} showed that LSTMs, GRUs and derivative methods show relatively similar performance for very short-term predictions on the time scale of a few hours.
 This timescale is where it is critical to accurately predict cloud movements.
 These methods have been taken as far as building genetic algorithms to generate the optimal neural network for the purpose \cite{jaidee_very_2019}.
-These genetic algorithms operate on the same principals as animal breeding.
+These genetic algorithms operate on the same principles as animal breeding.
 They generate multiple candidate neural networks and use the best networks as a base to generate new networks for the next generation.
-This is done for multiple generations and in the end the best neural network generated is used \cite{jaidee_very_2019}.\\
+This is done for multiple generations and in the end, the best neural network generated is used \cite{jaidee_very_2019}.\\
 Su et al.
@@
 The methods previously used all have in common that they must effectively predict the weather so they can predict power output.
-Here we explore whether better prediction accuracy can be reached
-if high quality, high resolution weather forecast data is used. This offloads the difficult task of predicting the weather from the responsibilities of the network. \\
-Can we construct a transformer based neural network which, by utilizing weather forecast data, is superior to neural networks which only utilize historical weather data?\\
+
+Here we explore using high-resolution weather forecast data from meteorological models, offloading a majority of this difficult task to the field of meteorology.
+Instead, we focus on correlating these forecasts with measurements and detecting and reducing temporal and spatial error in the general forecasts using newer methods like Temporal Fusion Transformers.
+What is the quality of the predictions this methodology produces? Can we create a model which is useful to power producers in operating their systems more efficiently?
+
+%Here we explore whether better prediction accuracy can be reached
+%if high quality, high resolution weather forecast data is used. This offloads a majority of the difficult task of predicting the weather from the responsibilities of the network. \\
+%Can we construct a transformer based neural network which, by utilizing weather forecast data, is superior to neural networks which only utilize historical weather data?\\
 

### 2022-04-08 methods.tex
A lot of work in methods

@@
 \section{Data}
-The data used comes from two primary sources
+This work relies on two types of data, the ground truth, the measured irradiance at the power station and the input, the irradiance values from high-resolution weather forecasts for the area.
+The data used spans the 425 days from August 1 2020 to September 30 2021 for the Ituverava solar power station run by Enel S.p.A.
+in Minas Gerais, Brazil.
+
 \subsection{Irradiance measurements}
+The irradiance was measured with seven Kipp \& Zonen CM21 pyranometers placed around the power station.
+The pyranometers have a reported maximum measurement error of 3\%.
+For input into the network, the highest and lowest measured values are dropped and the average value of the other five measurements is used.
+This data is provided with five-minute intervals, but the neural network operates hourly.
+To downsample the measurement data to an hourly format, we averaged the values for the twenty minutes before and after the top of each hour.
+Finally, the irradiance values were normalized linearly to the range 0-1 before input into the neural network.
+\newpage
+
 \subsection{Forecast Data}
-\subsubsection{Data Source}
-Text\cite{gueymard_global_2008, rognvaldsson_numerical_2013, jimenez_wrf-solar_2016}
-\subsubsection{Forecast Effectiveness}
+The weather forecast data is provided by Belgingur ehf., Reykjavik, Iceland.
+We use irradiance values from historical 24-hour weather forecasts generated by their WRF model, discussed in section~\ref{cha:WRF}.
+The forecasts have a temporal resolution of one hour and a spatial resolution of 2.5km.
+One of our models uses forecast data from a 3x3 grid centred on the solar power station, and the other model only the data from the centre of the grid.
+We use two irradiance values from the forecast, the downward irradiance and the downward diffuse irradiance.
+These irradiance values were normalized linearly to the range 0-1 before input into the neural network.
+
+\subsubsection{Data Source}\label{cha:WRF}
+\textcolor{red}{TODO: Talk about WRF} \cite{gueymard_global_2008, rognvaldsson_numerical_2013, jimenez_wrf-solar_2016}
+
 \subsection{Error metric}
+In other work done in this area, multiple error metrics are used.
+One error metric is used more commonly than others, Mean Absolute Percentage Error (MAPE) \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}, which is the reason that metric was chosen.
+When working with near-zero values MAPE becomes unhelpful as it is calculated by dividing by that near-zero value, so all cases where the irradiance is less than 15\% of the maximum value in the data set are dropped.
+Hence we do not evaluate early morning or late evening performance.
+While a valuable metric, MAPE does not  fully represent the value of the model.
+Other factors like confidence, discussed in section~\ref{sec:loss_function} and deviance from total irradiance for the day also represent key factors in the usefulness of the model.
+
+\newpage
 
 \section{Tools}
-
-For training models we used a computer with an AMD R9 5950X, 32GB RAM and an Nvidia Geforce RTX 3080Ti.
+The neural network was implemented in PyTorch 1.11 and PyTorch Forecasting 0.9.0.
+We used an NVIDIA GeForce RTX 3080Ti graphics card using CUDA for training.
+The computer had an AMD Ryzen 9 5950X CPU, 32GB of RAM and was running Windows 11.
 This allowed for quick training of models, as with the relatively small amount of data, training a model took 5-10 minutes.
 
+\newpage
 \section{Model}
     \subsection{Temporal Fusion Transformer}
-    Text \cite{lim_temporal_2020}
+    \textcolor{red}{TODO}
+    \cite{lim_temporal_2020}
+
+    \begin{figure}[ht!]
+        \centering
+        \includegraphics[scale=0.4]{imgs/TFT1.png}
+        \caption{An overview of the Temporal Fusion Transformer.\cite{lim_temporal_2020}
+        \label{fig:tft_overviwew}}
+    \end{figure}
+\newpage
+    \begin{figure}[ht!]
+        \centering
+        \includegraphics[scale=0.27]{imgs/TFT.png}
+        \caption{A detailed view of the architecture of the Temporal Fusion transformer.\cite{lim_temporal_2020}
+        \label{fig:ttgt_detail}}
+    \end{figure}
+
+    \newpage
     \subsection{Setup}
-    \subsubsection{Loss function}\label{cha:loss_function}
-    Quantile loss and loss quantiles
+    \textcolor{red}{TODO}
+
+        \begin{table}[ht!]
+        \begin{center}
+        \caption{The full model hyperparameters.
+        \label{tab:full_params}}
+        \vspace{0.5cm}
+        \begin{tabular}{|l|l|}
+        \hline
+        \textbf{Hyperparameter} & \textbf{Value} \\ \hline
+        Learning rate            & 0.01         \\ \hline
+        Batch size              & 24            \\ \hline
+        Dropout                 & 0.1            \\ \hline
+        Hidden size            & 64         \\ \hline
+        Hidden continuous size & 8         \\ \hline
+        Attention head size    & 1         \\ \hline
+        \end{tabular}
+
+        \end{center}
+        \end{table}
+
+        \begin{table}[ht!]
+        \begin{center}
+        \caption{The limited model hyperparameters.
+        \label{tab:lim_params}}
+        \vspace{0.5cm}
+        \begin{tabular}{|l|l|}
+        \hline
+        \textbf{Hyperparameter} & \textbf{Value} \\ \hline
+        Learning rate            & 0.005         \\ \hline
+        Batch size              & 24            \\ \hline
+        Dropout                 & 0.1            \\ \hline
+        Hidden size            & 40         \\ \hline
+        Hidden continuous size & 6         \\ \hline
+        Attention head size    & 1         \\ \hline
+        \end{tabular}
+
+        \end{center}
+        \end{table}
+
+    \subsubsection{Loss function}\label{sec:loss_function}
+    \textcolor{red}{TODO: Fix, find more sources and make sure this is correct}\\
+    We use the quantile loss  Temporal Fusion Transformers is quantile loss, which we use.
+    A model's prediction is a random value within a probability distribution \cite{}.
+If we make multiple predictions we can estimate this probability distribution by seeing within what range a selected proportion of the predictions fall \cite{koenker_quantile_2001}.
+PyTorch Forecasting's quantile loss function takes the Mean Absolute Error (MAE) of three quantiles, 90\%, 50\% and 10\%.
+[0.02, 0.1, 0.25, 0.5, 0.75, 0.9, 0.98] This gives a good picture of the model's confidence in a prediction.
+A smaller quantile range indicated more confidence.

### 2022-04-08 results.tex
A lot of work in methods

@@
 15.98\%, as well as better tracking in challenging conditions and a better ability to recognize uncertain conditions, generally reporting lower confidence when its predictions were wrong.
-Both models can be trained to achieve slightly lower MAPE, but that came at the cost of a higher tendency to underfit to variance, and higher false confidence.
+Both models can be trained to achieve slightly lower MAPE, but that came at the cost of a higher tendency to underfit variance, and higher false confidence.
 We show irradiance plots for select days representative of the models' performance in various conditions.
-These plots show the measured irradiance for the day, the models' predicted irradiance values, along with a 90\% confidence interval, detailed in section~\ref{cha:loss_function}.
+These plots show the measured irradiance for the day, the models' predicted irradiance values, along with a 90\% confidence interval, detailed in section~\ref{sec:loss_function}.
 The plots also show the forecast downward shortwave irradiance, the largest factor in measured irradiance, from the model input.
-Additionally the graphs contain the average MAPE for the day, and the sum of prediction and observation values.
+Additionally, the graphs contain the average MAPE for the day and the sum of prediction and observation values.
 The sum values act as a gauge for the total irradiance for the day.
@@
 \subsection{Low variance days}
-In the test dataset there are a number of clear days, where the sun was mostly or entirely unobstructed.
-The model predicted most of these days with very low error, in the range 2-5\%.
+In the test dataset, there are some clear days, where the sun was mostly or entirely unobstructed.
+The model predicted most of these days with very low error, in the range of 2-5\%.
 The 90\% confidence interval stayed within $\pm100 W/m^2$ for the whole day and the true prediction was almost spot-on, as can be seen in Fig.~\ref{fig:full_low_best}.
@@
 On days with some variance in the irradiance measured, the model was generally able to predict the irradiance with relatively good accuracy.
-The model has a tendency to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value was within the 90\% confidence interval.
+The model tends to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value was within the 90\% confidence interval.
 
@@
 \subsection{Low variance days}
-The model predicts most clear days with relatively high accuracy, but even on the best days it deviates from the quite a bit from the observed values, as shown in Fig.~\ref{fig:less_low_a}.
+The model predicts most clear days with relatively high accuracy, but even on the best days, it deviates quite a bit from the observed values, as shown in Fig.~\ref{fig:less_low_a}.
 The model sometimes under-predicts on clear days.
@@
 \end{figure}
+poopies manana
 
@@
 The model shows less accuracy when the irradiance is not steady.
-Predictions for most days have the smooth profile of an optimal day, but are scaled to fit closer to the lower total irradiance for the day.
+Predictions for most days have the smooth profile of an optimal day but are scaled to fit closer to the lower total irradiance for the day.
 The model is inconsistent in how well it tracks the observed irradiance and the confidence it does that with.

### 2022-04-25 introduction.tex
Implemented feedback from instructor

@@
 
-Solar power production is extremely dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016}.
-A single cloud can seriously impact the production of a solar farm.
-Such unreliability in a power production system is highly undesirable since production must match demand \cite{lee_forecasting_2018}.
+Solar power production is highly dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016}.
+A single cloud can adversely impact the production of a solar farm.
+Such unreliability in a power production system is undesirable since production must match demand \cite{lee_forecasting_2018}.
 When solar power production falls, the power demand it was serving must be fulfilled by other means.
 Fossil fueled power plants are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
-These plants often take hours to start, generally pollute more than other plants and are very expensive to run \cite{lee_forecasting_2018}.
+These plants often take hours to start, generally pollute more than other plants and are expensive to run \cite{lee_forecasting_2018}.
 Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}.
@@
 This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
-Recent advancements in deep learning have unlocked new tools that have been shown to be extremely effective in predicting the future behaviour of a system given enough relevant data.\\
+Recent advancements in deep learning have unlocked new tools that have been shown to be effective in predicting the future behaviour of a system given enough relevant data.\\
 
@@
 
+\textcolor{red}{TODO: Background on neural networks}\\
+
 \subsection{LSTMs and GRUs}
-\textcolor{blue}{Is there value in going deeper?}\\
 Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
@@
 
+The thesis is structured as follows.
+First we talked about the motivations and background in the introduction.
+Next we cover the methods, the data and tools used to achieve these goals.
+The proceeding chapter contains the results of the work, and finally we discuss the implications, strengths and flaws of the work, future work and end with conclusions.
+
 %Here we explore whether better prediction accuracy can be reached

### 2022-04-25 methods.tex
Implemented feedback from instructor

@@
 \section{Data}
-This work relies on two types of data, the ground truth, the measured irradiance at the power station and the input, the irradiance values from high-resolution weather forecasts for the area.
-The data used spans the 425 days from August 1 2020 to September 30 2021 for the Ituverava solar power station run by Enel S.p.A.
+This work relies on two types of data, \emph{the ground truth}, the measured irradiance at the power station and \emph{the input}, the irradiance values from high-resolution weather forecasts for the area.
+The data spans the 425 days from August 1st 2020 to September 30th 2021 for the Ituverava solar power station run by Enel S.p.A.
 in Minas Gerais, Brazil.
@@
 To downsample the measurement data to an hourly format, we averaged the values for the twenty minutes before and after the top of each hour.
-Finally, the irradiance values were normalized linearly to the range 0-1 before input into the neural network.
+Finally, the irradiance values were normalized linearly to the range $0.0-1.0$ before input into the neural network.
 \newpage
@@
 The weather forecast data is provided by Belgingur ehf., Reykjavik, Iceland.
-We use irradiance values from historical 24-hour weather forecasts generated by their WRF model, discussed in section~\ref{cha:WRF}.
+We use irradiance values from historical 24-hour weather forecasts generated by their WRF model, discussed in Section~\ref{cha:WRF}.
 The forecasts have a temporal resolution of one hour and a spatial resolution of 2.5km.
-One of our models uses forecast data from a 3x3 grid centred on the solar power station, and the other model only the data from the centre of the grid.
+One of our models uses forecast data from a 3x3 grid centred on the solar power station, and the other model the data from only the centre of the grid.
 We use two irradiance values from the forecast, the downward irradiance and the downward diffuse irradiance.
-These irradiance values were normalized linearly to the range 0-1 before input into the neural network.
+These irradiance values were normalized linearly to the range $0.0-1.0$ before input into the neural network.
 
@@
 While a valuable metric, MAPE does not  fully represent the value of the model.
-Other factors like confidence, discussed in section~\ref{sec:loss_function} and deviance from total irradiance for the day also represent key factors in the usefulness of the model.
+Other factors like confidence, discussed in Section~\ref{sec:loss_function} and deviance from total irradiance for the day also represent key factors in the usefulness of the model.
 
@@
         \vspace{0.5cm}
-        \begin{tabular}{|l|l|}
+        \begin{tabular}{|l|r|}
         \hline

### 2022-04-25 results.tex
Implemented feedback from instructor

@@
 \chapter{Results}
-Both models showed a strong ability to predict solar irradiance one day into the future.
-The full model, using weather forecast data from the surrounding area, showed both a lower Mean Absolute Percentage Error (MAPE), at 13.21\% vs.
-15.98\%, as well as better tracking in challenging conditions and a better ability to recognize uncertain conditions, generally reporting lower confidence when its predictions were wrong.
-Both models can be trained to achieve slightly lower MAPE, but that came at the cost of a higher tendency to underfit variance, and higher false confidence.
+Both models show a strong ability to predict solar irradiance one day into the future.
+The full model, using weather forecast data from the surrounding area, shows both a lower Mean Absolute Percentage Error (MAPE), at 13.21\% vs.
+15.98\%, as well as better tracking in challenging conditions and a better ability to recognize uncertain conditions, generally reporting lower confidence when its predictions are wrong.
+Both models can be trained to achieve slightly lower MAPE, but that comes at the cost of a higher tendency to underfit variance, and higher false confidence.
 We show irradiance plots for select days representative of the models' performance in various conditions.
@@
 \section{Full model}
-The full model, utilizing irradiance data from surrounding areas achieved an average MAPE of 13.21\% on the test dataset. The maximum daily average MAPE was 138.87\% and the minimum was 2.50\%. 
+The full model, utilizing irradiance data from surrounding areas achieves an average MAPE of 13.21\% on the test dataset.
+The maximum daily average MAPE is 138.87\% and the minimum is 2.50\%.
+Figure~\ref{fig:days_full} shows the average MAPE the full model achieves for the test dataset on a day-by-day basis.
 \begin{figure}[ht!]
@@
 In the test dataset, there are some clear days, where the sun was mostly or entirely unobstructed.
-The model predicted most of these days with very low error, in the range of 2-5\%.
-The 90\% confidence interval stayed within $\pm100 W/m^2$ for the whole day and the true prediction was almost spot-on, as can be seen in Fig.~\ref{fig:full_low_best}.
+The model predicts most of these days with very low error, in the range of 2-5\%.
+The 90\% confidence interval stays within $\pm100 W/m^2$ for the whole day and the true prediction is almost spot-on, as can be seen in Figure~\ref{fig:full_low_best}.
 \begin{figure}[ht!]
@@
     \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/best_2.png}}\qquad
-    \caption{Examples of clear days where the model preformed very well.
+    \caption{Examples of clear days where the model preforms very well.
     \label{fig:full_low_best}}
@@
 
-The model did not manage near-perfect predictions on all of these clear days, and made a few under-predictions, as can be seen in Fig.~\ref{fig:full_low_worst}.
-These predictions were still even and within the 90\% confidence interval, which was slightly larger than for the other clear day predictions. These predictions had a higher MAPE of 10-13\%.
+The model does not manage near-perfect predictions on all of these clear days, and makes a few under-predictions, as can be seen in Figure~\ref{fig:full_low_worst}.
+These predictions still predict steady sunshine, and are within the 90\% confidence interval, which is slightly larger than for the other clear day predictions.
+These predictions have a higher MAPE of 10-13\%.
 \begin{figure}[ht!]
@@
     \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/worst_2.png}}\qquad
-    \caption{Examples of clear days where the model did not perform well.
+    \caption{Examples of clear days where the model does not perform well.
     \label{fig:full_low_worst}}
@@
 \subsection{Medium variance days}
-On days with some variance in the irradiance measured, the model was generally able to predict the irradiance with relatively good accuracy.
-The model tends to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value was within the 90\% confidence interval.
+On days with some variance in the irradiance measured, the model is generally able to predict the irradiance with relatively good accuracy.
+The model tends to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value is within the 90\% confidence interval.
 
-Fig.~\ref{fig:full_med_best} shows examples of good predictions with high confidence, though the confidence is generally lower the farther the irradiance is from the value of a clear day.
+Figure~\ref{fig:full_med_best} shows examples of good predictions with high confidence, though the confidence is generally lower the farther the irradiance is from the value of a clear day.
 \begin{figure}[ht!]
@@
     \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_g_4.png}}\qquad
-    \caption{Examples of days with some drops in irradiance, where the model performed very well.
+    \caption{Examples of days with some drops in irradiance, where the model performs very well.
     \label{fig:full_med_best}}
@@
 
-In Fig.~\ref{fig:full_med_med} we can see examples of days where the model achieves higher error, MAPE in the range 19-35\%, but with generally larger 90\% confidence intervals so the model still keeps the measured values close to or within them.
+In Figure~\ref{fig:full_med_med} we see examples of days where the model achieves higher error, MAPE in the range 19-35\%, but with generally larger 90\% confidence intervals so the model still keeps the measured values close to or within them.
 \begin{figure}[ht!]
@@
     \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_m_5.png}}\qquad
-    \caption{Examples of days with some drops in irradiance, where the model performed somewhat well.
+    \caption{Examples of days with some drops in irradiance, where the model performs somewhat well.
     \label{fig:full_med_med}}
@@
 
-For a few medium variance days, the model did not track well.
-Figs.~\ref{fig:full_med_bad_a} and \ref{fig:full_med_bad_d} show where the model predicted a constant lower irradiance when half the day had high irradiance and the other half low.
-Figs.~\ref{fig:full_med_bad_b} and \ref{fig:full_med_bad_c} show the model predicting a mostly or fully  sunny day when it wasn't.
-In all of these cases, the observed irradiance was far outside the 90\% confidence interval.
+For a few medium variance days, the model does not track well.
+Figures~\ref{fig:full_med_bad_a} and \ref{fig:full_med_bad_d} show where the model predicts a constant lower irradiance when half the day had high irradiance and the other half low.
+Figures~\ref{fig:full_med_bad_b} and \ref{fig:full_med_bad_c} show the model predicting a mostly or fully sunny day when it wasn't.
+In all of these cases, the observed irradiance is far outside the 90\% confidence interval.
 \begin{figure}[ht!]
@@
     \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/medium/med_b_4.png}\label{fig:full_med_bad_d}}\qquad
-    \caption{Examples of days with some drops in irradiance, where the model performed badly.
+    \caption{Examples of days with some drops in irradiance, where the model performs badly.
     \label{fig:full_med_bad}}
@@
 \subsection{High variance days}
-A few days in the dataset had very low irradiance for part of, or the whole day or large fluctuations in irradiance.
-These situations were most difficult for the model to predict.
-Figs.~\ref{fig:full_high_a}, \ref{fig:full_high_b} and \ref{fig:full_high_c} show the model somewhat following the observations, but with a high MAPE of 38-80\% because of offsets in time and/or magnitude.
-The observed value was often outside the 90\% confidence interval of the predictions.
-For Figs.~\ref{fig:full_high_a}, \ref{fig:full_high_b} the sum of predicted values is close to the sum of observed values, indicating relatively high accuracy for the days taken as a whole.
-Fig.\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that receives very little irradiance.
+A few days in the dataset had very low irradiance for part of, or the whole day, or large fluctuations in irradiance.
+These situations are the most difficult for the model to predict.
+Figures~\ref{fig:full_high_a}, \ref{fig:full_high_b} and \ref{fig:full_high_c} show the model somewhat following the observations, but with a high MAPE of 38-80\% because of offsets in time and/or magnitude.
+The observed value is often outside the 90\% confidence interval of the predictions.
+For Figures~\ref{fig:full_high_a}, \ref{fig:full_high_b} the sum of predicted values is close to the sum of observed values, indicating relatively high accuracy for the days taken as a whole.
+Figure\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that received very little irradiance.
 \begin{figure}[ht!]
@@
 \section{Less forecast data}
-The limited model, utilizing irradiance data only from the one forecast region, and not surrounding regions, achieved an average MAPE of 15.98\% on the test dataset.
-The maximum daily average MAPE was 122.87\% and the minimum was 3.47\%.
+To confirm that the full model utilizes the input data for the surrounding area, we also trained the model on only irradiance forecast for the cell the power station is located in.
+The limited model achieves an average MAPE of 15.98\% on the test dataset.
+The maximum daily average MAPE is 122.87\% and the minimum is 3.47\%.
+Figure~\ref{fig:days_less} shows the average MAPE the limited model achieved for the test dataset on a day-by-day basis.
 \begin{figure}[ht!]
@@
 \subsection{Low variance days}
-The model predicts most clear days with relatively high accuracy, but even on the best days, it deviates quite a bit from the observed values, as shown in Fig.~\ref{fig:less_low_a}.
+The model predicts most clear days with relatively high accuracy, but even on the best days, it deviates quite a bit from the observed values, as shown in Figure~\ref{fig:less_low_a}.
 The model sometimes under-predicts on clear days.
-In Fig.~\ref{fig:less_low_b} we can see an example of a day where, even though the 90\% confidence area is large, the observed irradiance is outside it.
+In Figure~\ref{fig:less_low_b} we can see an example of a day where, even though the 90\% confidence area is large, the observed irradiance is outside it.
 \begin{figure}[ht!]
@@
 The model is inconsistent in how well it tracks the observed irradiance and the confidence it does that with.
-Sometimes, as shown in Fig.~\ref{fig:less_med_a} the model tracks relatively well, with the observations mostly within the rather tight 90\% confidence interval of roughly $+50/-150 W/m^2$.
-In Fig.~\ref{fig:less_med_b} we see an example where the model was wrong with very high confidence for the first half of the day.
-Fig.~\ref{fig:less_med_c} shows the opposite, the model predicts low irradiance for the whole day with a very large 90\% confidence interval of roughly $\pm200 W/m^2$.
+Sometimes, as shown in Figure~\ref{fig:less_med_a} the model tracks relatively well, with the observations mostly within the rather tight 90\% confidence interval of roughly $+50/-150 W/m^2$.
+In Figure~\ref{fig:less_med_b} we see an example where the model is wrong with very high confidence for the first half of the day.
+Figure~\ref{fig:less_med_c} shows the opposite, the model predicts low irradiance for the whole day with a very large 90\% confidence interval of roughly $\pm200 W/m^2$.
 \begin{figure}[ht!]
@@
 \subsection{High variance days}
-For the few days in the dataset that had very low irradiance for part of, or the whole day or large fluctuations in irradiance, the model showed high error, MAPE in the range 53-123\% with varying confidence.
-Figs.~\ref{fig:less_high_a} and \ref{fig:less_high_b} show large over-predictions with high confidence, and Figs.~\ref{fig:less_high_c} and \ref{fig:less_high_d} show days where the 90\% confidence interval covered the top 60\% of the graph, yet it did not cover the observations.
+For the few days in the dataset that had very low irradiance for part of, or the whole day or large fluctuations in irradiance, the model displays high error, MAPE in the range 53-123\% with varying confidence.
+Figures~\ref{fig:less_high_a} and \ref{fig:less_high_b} show large over-predictions with high confidence, and Figures~\ref{fig:less_high_c} and \ref{fig:less_high_d} show days where the 90\% confidence interval covers the top 60\% of the graph, yet it does not cover the observations.
 \begin{figure}[ht!]

### 2022-05-06 MSC-GUDMUNDSSON-2022.tex
Added content to methods, added abstract, cleanup, new diagrams and more.

%% ---------------------------------------------------------------
%% $URL: https://repository.cs.ru.is/svn/thesis-template/trunk/ruthesis/latex/DEGREE-NAME-YEAR.tex $
%% $Id: DEGREE-NAME-YEAR.tex 360 2019-02-13 22:04:35Z foley $
%% This is a template LaTeX file for dissertations, theses, or reports at Reykjavík University
%% 
%% Comments and questions can be sent to the RU LaTeX group (latex AT list.ru.is) 
%% ---------------------------------------------------------------

%% METHOD:
%% 0) Read ruthesis/thesis-instructions.pdf
%%    If it is missing, goto https://repository.cs.ru.is/svn/thesis-template/trunk/ruthesis/thesis-instructions.pdf
%% 0.2) Subscribe to the announcements email list at
%%    https://list.ru.is/mailman/listinfo/latex-announcements
%% 1 LaTeX instructions.tex or goto http://afs.rnd.ru.is/project/thesis-template/trunk/ruthesis/latex/instructions.pdf
%% 2) Copy the template files (or unzip) to your working area
%% 3) Rename this file (if needed) with your information e.g. MSC-FOLEY-2007.tex
%% 4) Modify this file to fit your needs (please follow all comments below in the text)
%% 5) For making bibliographies, run "biber".  You can also change
%%    this back to "bibtex".  See below in "Bibliography options".

%%%%%%% CHOOSE ONE OF THESE %%%%%%%%%%%%%%%
%% projectreport: Project report (CS)
%% bachelors: Bachelor of Science thesis
%% masters: Master of Science thesis
%% doctorate: Doctor of Philosophy dissertation
%
%%%%%%% CHOOSE ONE OF THESE %%%%%%
%% 
%% draft: speed up processing by skipping graphics and adding useful
%%     information for editing.  Also sets spacing to double so that it is easier to
%%     write editing marks on paper copy.
%% proof:  proofreading version (final formatting with warnings)
%% final: generate document for submission, removing FIXMEs, and
%%     other markup.  Throw error if any fatal FIXMEs still in document.
%%
%%%%%%% CHOOSE ONE OF THESE IF APPLICABLE %%%%%%
%%
%% deptsse: School of Science and Engineering
%% deptscs: School of Computer Science
%%
%%%%%%%% CHOOSE ANY COMBINATION OF THESE %%%%%%%%%%%%
%%
%% forcegraphics: force graphics, etc. to be included, even in draft mode
%% debug:  writes more messages to the log file, adds debugging output 
%%     and sizing boxes
%% icelandic: thesis is in Icelandic
%% oldstyle:  use the PhD headers and footers from the old CS template
%% online: for online versions (skip blank pages)
\documentclass[online,masters,deptscs,forcegraphics,final]{ruthesis}

%%%%%%%%%%%%%%%%%%%% TeXStudio Magic Comments %%%%%%%%%%%%%%%%%%%%%
%% These comments that start with "!TeX" modify the way TeXStudio works
%% For details see http://texstudio.sourceforge.neit/manual/current/usermanual_en.html   Section 4.10
%%
%% What encoding is the file in?
% !TeX encoding = UTF-8
%% What language should it be spellchecked?
% !TeX spellcheck = en_US
%% What program should I compile this document with?
% !TeX program = xelatex

%%%%%%%%%%%%%%%%%%%% Bibliography options %%%%%%%%%%%%%%%%%%%%%
%% We suggest switching from bibtex to biblatex/biber because it is better able
%% to deal with Icelandic characters and other bibliography issues
%% As long as you use biblatex instead of bibtex by itself, it will at least
%%  generate a document without errors.
%% !!!If you are using TeXStudio, don't forget to update the bibliography setting!!!
\usepackage[backend=biber,bibencoding=utf8,style=ieee]{biblatex}
%\DeclareLanguageMapping{american}{american-apa}  
% need to declare mapping for style=apa to alphabetize properly
% If you set backend=bibtex, it will use bibtex for processing (old way)
%    this can work with Icelandic characters, but you may get weird results.
%    bibtex does not know how to sort Þ and ð
% if you set backend=biber, you can use UTF8 characters such as Þ and
%     ð  but you will have to remember to switch from using bibtex to 
%     biber in your client
% If you use JabRef, make sure the file is encoded in UTF-8 which is
%    not the default.

%% This tells TeXStudio to use biber
% !TeX TXS-program:bibliography = txs:///biber
%% This also sets the bibliography program for TeXShop and TeXWorks
% !BIB program = biber

% Where is your reference library?
\addbibresource{references.bib}
\addbibresource{more_references.bib}

%%%%%%%%%%%%%%%%%%% CUSTOMIZATIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% It is not recommended that you customize this file nor
%% ruthesis.cls.  Just fill in the necessary fields.  You should put
%% your macros and packages into a separate file so that it is easier
%% to use updates to the template.  The custom.sty file was created
%% for this reason.  We load this much later so that it can overrite
%% any existing settings
\IfFileExists{custom.sty}{\usepackage{custom}}{}

%%%%%%%%%%%%%%% INFORMATION %%%%%%%%%%%%%%%%%%5
%% University information must be multilingual to deal with the
%%  required cover pages and abstract on thesis
%% NOTE: This may not be required for other reports!!!

%% Babel Icelandic macros are setup  on RedHat at
%% /usr/share/texlive/texmf-dist/tex/generic/babel/icelandic.sty
%% /usr/share/texlive/texmf-dist/tex/generic/babel-icelandic/icelandic.ldf

%% Multilingual macros
%\newML{macroname}{englishword}{icelandicword}
%  creates \macronameML
%    \MLmacroname[english] - returns the english word
%    \MLmacroname[icelandic] - returns the icelandic word
%    \MLmacroname  - uses the current language setting
% Some useful ones have already been defined, but can be redefined
%% Predefined: \MLIceland \MLReykjavikUniversity \MLUniversityIceland

%% What institute?  Default is RU.
%\setInstitution{\MLReykjavikUniversity}
% \newML{InstitutionAddress}{Menntavegur 1\\101 Reykjavík, Iceland}
% {Menntavegi 1\\101 Reykjavík, Ísland}
% \setInstitutionAddress{\MLInstitutionAddress}
% \newML{Tel}{Tel.}{Sími}
% \setInstitutionPhone{\MLTel{} +354 599 6200\\
% Fax +354 599 6201}
% \setInstitutionURL{www.ru.is}

%% ONLY SET DEPARTMENT IF YOU HAVE NOT USED THE deptsse or deptscs OPTION!
%% Department and degree program
%\newML{ND}{New Department}{Nytt deild}
%\setSchool{\MLND}

%% Set your program of study
\newML{program}{Software Engineering}{hugbúnaðarverkfræði}
\program{\MLprogram}

%% Degree long name.  If not already defined, you can create a macro
%\newML{DEGREE}{English Degree Name}{Icelandic Degree Name}
%% Default is set based upon doctorate vs masters option
%% Predefined: \MLMSc \MLPhd
%\setDegreelong{\MLMSc}

%% Degree abb, change if default is not right
%% Default is set based upon doctorate vs masters option
%\degreeabbrv{Sc.D.}

%\setFrontLogo{reyst-logo}
%% Use this if you need a different front logo on the first page
%% e.g. reyst-logo

%% Date in english and icelandic
%% NOTE: THIS IS THE DATE OF THE SUPERVISOR'S SIGNATURE!!!!!!
%% Predefined: \MLjan, \MLfeb, \MLmar, ... \MLdec
%\whensigned{day}{month}{year} %day is only used on some formats, but you must put something.
\whensigned{10}{\MLapr}{2022}

%% Title first in English then Icelandic
%% You need to put both a normal case and ALL CAPS version into the macros.
%%
\newML{Title}{Forecasting Solar Power Production Using Deep Learning and High Resolution Weather Forecasts}{Sólarorkuframleiðsluspár með djúpnámi út frá háupplausnar veðurspám}
\newML{TITLE}{FORECASTING SOLAR POWER PRODUCTION USING DEEP LEARNING AND HIGH RESOLUTION WEATHER FORECASTS}{SÓLARORKUFRAMLEIÐSLUSPÁR MEÐ DJÚPNÁMI ÚT FRÁ HÁUPPLAUSNAR VEÐURSPÁM}
%%
\setTitle{\MLTitle}{\MLTITLE}
%% ***** Special Titles ******
%% If the title must be formatted specifically for the cover page or internal pages
%% (typically via line-breaks using the \newline command) then the following commands must be used 
%%
%\setTitleCover{\MLTITLE}
%% These two for the internal cover pages, usually not needed
%\newML{TitleInternal}{Internal Title}{Icelandic Internal Title}
%\setTitleInternal{\MLTitleInternal}

%% Author name (should be the same in any language, if not use \newML)
%% If you are writing a Project report with multiple authors, separate them with \\:
%% To keep the names typeset together, you want to use non-breaking spaces: ~
%\author{Firstname1~Lastname1\\Firstname2~Lastname2}
\author{Einar~Helgi~Guðmundsson}

%% If the name must be formatted specifically for the signature page
%% (typically via line-breaks) then the following command must be used 
%\setAuthorSignature{Student\\Name}
%% This macro adjusts the author name in the headers of the oldstyle formatting
%\setAuthorHeader{StudentLast}

%%% TODO:  Move the bachelor's form separately -- it confuses people. --foley
%%%%%%%%%%%%%%%%%%%%%%%%%% Project Report or Bachelor's Only!!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setCourse{VT LOK 1012}

%%%%%%%%%%%%%%%%%%%%%%%%%% Bachelors Only!!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setID{160992--2989}%kennitala
\setSemester{2022-2}
\setShortSignedDate{10.4.2022}

\setOrganization{Belgingur ehf.\\Grensásvegur 9\\108 Reykjavík}
\setSubProgram{Hugbúnaðarverkfræði}

%% If the thesis is confidential, uncomment this with the date it can be released
%\setClosedDistribution{10.1.2016}%

%% Put your keywords here in English, then Icelandic.  Separate them with commas.
\newML{keywords}{Deep Learning, Solar Power, Forecast}{Djúpnám, Sólarorka, spá}
\setKeywords{\MLkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Masters Only!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% How many credits (ECTS) on Master's degree
%% Usually 30 or 60
\ects{60}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Doctorate Only!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some Computer Science Thesis have an ISSN number.
%% Most other documents do not.
%\bookidnumber{ISSN: 1670-8539} 
%% ID numbers are optional, but nice for sorting in libraries

%% International Standard Book Number (ISBN)
%% This is what most people should use if the thesis is being published.

%% International Standard Serial Number (ISSN)
%% This is usually only for a PhD dissertation as part of a series when published
%%   Computer Science: 1670-8539

%% Additional degrees?  (optional, usually not needed)
%\adddegree{(list of degrees in appendix)}{(sjá lista yfir prófgraður í viðauka)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% List the entire committee.  Each member has a name (degree should be omitted, unless it is not PhD),
%% Supervisor(s) must appear first
%% On a Bachelors, there is usually only one supervisor and one examiner.

%% Format for each entry:
%%  \personinfo{Name}{Role}{Job Title}{Company/institution}{Country}
%% Predefined macros: \MLSupervisor \MLSupervisors \MLExaminer \MLExaminers

%% Change these to singular/plural as needed.
%% Just uncomment and change the plurality of the macro.
%\setSupervisorHeading{\MLSupervisors}
%\setExaminerHeading{\MLExaminer}

%% Predefined macros:
%% \MLSeniorProfessor \MLProfessor \MLAssociateProfessor \MLAdjunctProfessor \MLEmeritusProfessor \Iceland
%% \MLReykjavikUniversity \MLUniversityIceland

%% Bachelors: primary advisor (Umsjónarkennari), ONLY ONE!
%% All others: As many as you want
\supervisors{
  \personinfo{Dr. Yngvi Björnsson, PhD}{\MLSupervisor}{\MLProfessor}{\MLReykjavikUniversity}{\MLIceland}
  \personinfo{Dr. Ólafur Rögnvaldsson, PhD}{Examiner}{Managing Director}{Belgingur ehf.}{\MLIceland}
%  \personinfo{Ian M. Great}{Co-advisor}{\MLProfessor}{Hochschule Düsseldorf}{Germany}
}

%% Bachelors: secondary advisor (Leiðbeinandi), ONLY ONE
%% All others: As many as you want
\examiners{
  \personinfo{Tough E. Questions}{\MLExaminer}{Associate Professor}{Massachusetts Institute of Technology}{USA}

}

%% An abstract is required to be in both Icelandic and English for most degrees.
%% It is considered good form to limit the abstract to a single paragraph in each language,
%%   at 300 words.  Refer to your degree's instructions.
%% Note: Icelandic quotation marks cannot be typeset using "` and "'.  You should use \enquote{}
%% this is probably due to interactions with the MultiLingual macros.
%% TODO: turn this into more sensible macros to avoid confusion --foley
\newML{AbstractText}{
The world is moving at increasing speed away from generating electricity using fossil fuels, toward more environmentally friendly options, such as solar power. Solar power production is volatile and highly dependent on weather conditions, primarily cloud cover. Instability and unpredictability are undesirable in power production since production must match demand. In order to ensure power availability, highly polluting and expensive fossil power stations, \enquote{peaker plants}, are kept on standby to react to the instability in solar power. The ability to reliably predict solar power production provides more predictability to power grid operators, allowing for greater efficiency in grid operation. 
Valuable data exists in weather forecasts, which can be harnessed to increase predictability in solar power generation. We use novel deep learning methods, attention, and specifically Temporal Fusion Transformers, to forecast the irradiance at a solar power station, which correlates very strongly with power output. We base this on Belgingur ehf.'s high-resolution weather forecast data, the 24-hour forecast irradiance values for the area of the power station.
Our model achieves a very low average Mean Absolute Percentage Error of 13.21\% on the test dataset for one-day forecasts. In addition, the model outputs a confidence level for its predictions providing insight to a grid operator on whether to expect unpredictability in the power output. The model is able to generate more accurate predictions when utilizing forecast data from a wider area around the power station than exclusively data from the power station‘s immediate vicinity.

}  
{
Íslenska!
} % Icelandic abstract goes here
\setAbstract{\MLAbstractText}

%%%%%%%%%%%%%%INDEX SETUP %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Indexes, and other auto-generated material
%% The Memoir package (which we use) automatically generates the index
%% See section 17.2 on page 302 of the guide
%% http://texdoc.net/texmf-dist/doc/latex/memoir/memman.pdf
%% This means you have to run "makeindex DEGREE-NAME-YEAR"
%% !!!Do not load any of the index packages, they cause problems with Memoir!!!
%% !!!You have been warned!!!
%% Note that memoir changes the [] options to only be for filenames, not other options!
\makeindex{}
\indexintoc{}

%% For abbreviations, you may want to try
%% Watch out though, each new index writes another external file and 
%% latex can only write a limited number of them
%%\usepackage[intoc]{nomencl} % intoc: In Table of Contents
%% remember to run:
%% makeindex filename.nlo  -s nomencl.ist -o filename.nls

\finalifforcegraphics{hyperref} %hyperlinks even in draft mode
\usepackage[hidelinks]{hyperref} 
%% !!!Must be the last package loaded except otherwise mentioned!!!!
%% \usepackage{hypcap}  %% puts link at top of figure, must be after hyperref
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{afterpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% DOCUMENT START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%% Some elements have different names on the RU Masters rules
%% They will be annotated with RUM: "name"
\frontmatter{} % setup formatting at beginning

%\frontcover{}%%If you want to see what it looks like with the printed cover
%% TODO:  link to fill-in PDF file on RU website

\frontrequiredpages{}%% the various signaturepages and abstract
%%% WARNING:  if you get an error on the previous line, it is probably because
%%% you put a bad macro or something strange in a title, author, or abstract.

\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
  have not already done so, \LaTeX{} the \path{instructions.tex} to
  learn how to setup your document and use some of the features.  You
  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
  There is also more information on working with \LaTeX{} at
  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
  This includes common problems and fixes.

  This page will disappear in anything other than draft mode.}{}

%% Dedication is optional, comment out if it is absent
%% RUM: Not mentioned
\begin{dedications}
  I dedicate this to my spouse/child/pet/power animal.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
% RUM: "Acknowledgements (optional)"
\coverchapter{Acknowledgements} 
\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}\sourceatright{Douglas Adams\cite{adams84fish}}
\vspace{\baselineskip}

\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
  Note that it is important to acknowledge any funding that helped in the work}

The project is partially funded by the Icelandic Technology Development Fund through grant nr. 1910034-0611.

\coverchapter{Preface}
% RUM: "Preface (optional)"
This dissertation is original work by the author, Einar~Helgi~Guðmundsson.

\draftnote{The preface is an optional element
  explaining a little who performed what work.  See
  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
  for suggestions.

  List of publications as part of the preface is
  optional unless elements of the work have already been published.
  It should be a comprehensive list of all publications in which
  material in the thesis has appeared, preferably with references to
  sections as appropriate.  This is also a good place to state
  contribution of student and contribution of others to the work
  represented in the thesis.}

%\coverchapter{Publications}
%% RUM: Not mentioned, this was found in the CS thesis template.  
%% Maybe more applicable to PhD dissertations?
%%% Probably a duplication from before Preface became standard.

\starttables{}% setup formatting
%% TOC, list of figures and list of tables are required
\tableofcontents{}\clearpage%%RUM: "Table of contents"
\listoffigures{}\clearpage%%RUM: "List of figures"
\listoftables{}\clearpage%%RUM: "List of tables"

%\coverchapter{List of drawings and enclosed material}
%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"

\listoffixmes{}
% if using fixme package, lists what needs to be done

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
PhD &Doctor of Philosophy\\
ANN &Artificial Neural Network\\
NN &Neural Network\\
RNN &Recurrent Neural Network\\
GRU &Gated Recurrent Unit\\
LSTM &Long Short Term Memory\\
NLP &Natural Language Processing\\
TCNN &Temporal Convolutional Neural Network\\
NARXNN &Non-linear Auto Regressive Neural Network\\
TFT &Temporal Fusion Transformer\\
GRN &Gated Residual Network\\
MAPE &Mean Absolute Percentage Error\\
MAE &Mean Absolute Error\\
CUDA &Compute Unified Device Architecture\\
AMD &Advanced Micro Devices inc.\\
CPU &Central Processing Unit\\
RAM &Random Access Memory\\

\end{tabular}

%\coverchapter{List of Symbols}%%RUM: Not mentioned
%\begin{tabular}{lll}
%Symbol &Description &Value/Units\\
%$E$ &Energy &\si{\joule}\\
%$m$ &Mass &\si{gram}\\
%$c$ &Speed of Light &\SI{2.99E8}{\meter\per\second\square}\\
%\end{tabular}

%% This command prepares for the actual text, e.g. by 
%% calling \mainmatter{}
\starttext{}

%% ---------------------------------------------------------------
%% From this point on, it is standard Latex, except the very end.
%% This is a "report"-based template, so the top-level heading 
%% is \chapter{}

%% WARNING: Make sure that all of these files (and any new ones)
%% are UTF-8 otherwise you will get weird encoding errors.

%% The default division is IMRAD, you may want to divide differently
%% See the introduction for guidance.

\include{introduction}%%RUM: Introduction
\include{methods}%%RUM: "Methods"
\include{results}%%RUM: "Results"
\include{discussion}%%RUM: "Discussion"

%% ---------------------------------------------------------------
\printbibliography{} %%RUM: "References"

%% If appendices are needed, uncomment the following line
%% and include the appendices in separate files
%\appendix{}%%RUM: "Appendicies (as appropriate)
%\input{code} % as an example, perhaps some of your code

%\backmatter{} % Sections after this don't get numbers
%% We prefer that all elements be numbered

%%%%%%%%%%%%% SHOW INDEX %%%%%%%%%%%%%%%%%%
%% Index, optional.  A good idea on longer documents

% You can put instructions at the beginning of the index:
%\renewcommand{\preindexhook}{%
%  The first page number is usually, but not always,
%  the primary reference to the indexed topic.\vskip\onelineskip}

%% You may have to run "makeindex <FILENAME>" to have it be generated
%% Depending upon which package you chose.
%% 
\clearforchapter{}
\printindex{}%%RUM: Not mentioned

%\backcover{}%%RUM: "Back cover (only Phd)
\end{document}

%% ---------------------------------------------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: xetex
%%% End:

### 2022-05-06 introduction.tex
Added content to methods, added abstract, cleanup, new diagrams and more.

@@
 \textcolor{red}{TODO: Background on neural networks}\\
+Artificial Neural Networks (ANNs), generally known as just Neural Networks (NNs) were first theorized in the 1940s \cite{citation needed}, though it was much later, in the\\
+NN function\\
 
-\subsection{LSTMs and GRUs}
+\subsection{LSTMs and GRUs\label{cha:RNN}}
 Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
@@
 
-\subsection{Transformers}
+\subsection{Transformers\label{cha:transformer}}
 %Hard to train
@@
     \centering
-    \includegraphics[scale=0.7]{imgs/transformer.png}
+    \includegraphics[scale=0.95]{imgs/transformer.png}
     \caption{The transformer architecture.\cite{vaswani_attention_2017}
@@
     \includegraphics[scale=0.5]{imgs/encoder_decoder.jpeg}
-    \caption{The encoder/decoder structure
+    \caption{The encoder/decoder structure \cite{nechu_what_2020}
     \label{fig:encoder_decoder}}
@@
 A single attention head can put all the focus on a single element.
-Multiple attention heads can be used to let the transformer model consider the context of more elements at the same time.\cite{rohrer_transformers_2021}
+Multiple attention heads can be used to let the transformer model consider the context of more elements at the same time \cite{rohrer_transformers_2021}.
 

### 2022-05-06 methods.tex
Added content to methods, added abstract, cleanup, new diagrams and more.

@@
 Finally, the irradiance values were normalized linearly to the range $0.0-1.0$ before input into the neural network.
-\newpage
 
-\subsection{Forecast Data}
+\subsection{Forecast Data\label{cha:forecast_data}}
 The weather forecast data is provided by Belgingur ehf., Reykjavik, Iceland.
@@
 
-\newpage
-
 \section{Tools}
@@
 
-\newpage
 \section{Model}
     \subsection{Temporal Fusion Transformer}
-    \textcolor{red}{TODO}
-    \cite{lim_temporal_2020}
+    \textcolor{red}{TODO: more}
+
+    We utilize a Temporal Fusion Transformer (TFT) model for the irradiance prediction.
+    TFTs are  general networks which use attention at their core to generate multi-horizon forecasts with the ability to consider
+    known information about the future, other related time series, and static metadata for the forecast \cite{lim_temporal_2020}. Figure~\ref{fig:tft_overviwew} shows an overview of how TFTs operate.
+
+    \afterpage{%
+    \begin{figure}
+        \centering
+         \makebox[\textwidth][c]{\includegraphics[scale=0.5]{imgs/TFT_overview.jpg}}
+        \caption{An overview of the Temporal Fusion Transformer.
+        \label{fig:tft_overviwew}}
+    \end{figure}
+    \clearpage
+    }
+    TFTs are composed of various components, each primarily addressing a single input type (i.e. static, known or observed inputs), aimed at maximizing the model versatility across various problems.
+
+    \textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data.\\
+    A Gated Residual Network (GRN), shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}.\\
+    \indent\textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.\\
+    \indent\textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
+    \indent\textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
+    \indent\textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, show a confidence interval for the predictions \cite{lim_temporal_2020}. \\
+    Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.
+
+    \afterpage{%
+    \begin{figure}
+        \centering
+         \makebox[\textwidth][c]{\includegraphics[scale=0.37]{imgs/TFT.png}}
+        \caption{A detailed view of the architecture of the Temporal Fusion transformer.\cite{lim_temporal_2020}
+        \label{fig:ttgt_detail}}
+    \end{figure}
+    \clearpage
+    }
+
+    \subsection{Setup}
+    \textcolor{red}{TODO: finish --- early stopping}\\
+    We trained two models.
+One which utilizes weather forecast data from an area surrounding the power station, and one which does not, discussed in Section~\ref{cha:forecast_data}.
+The limited model is intended to show that the model is able to glean additional information from the surrounding area data.
 
@@
         \centering
-        \includegraphics[scale=0.4]{imgs/TFT1.png}
-        \caption{An overview of the Temporal Fusion Transformer.\cite{lim_temporal_2020}
-        \label{fig:tft_overviwew}}
-    \end{figure}
-\newpage
-    \begin{figure}[ht!]
-        \centering
-        \includegraphics[scale=0.27]{imgs/TFT.png}
-        \caption{A detailed view of the architecture of the Temporal Fusion transformer.\cite{lim_temporal_2020}
-        \label{fig:ttgt_detail}}
+        \includegraphics[scale=0.6]{imgs/solar_TFT_overview.jpg}
+        \caption{An overview of  the setup of the neural network.
+        \label{fig:solar_tft_overviwew}}
     \end{figure}
 
-    \newpage
-    \subsection{Setup}
-    \textcolor{red}{TODO}
+    \subsubsection{Inputs}
+        The dataset was split into a training set, composing 70\% of the data, a validation set, composing 10\% of the data and finally 20\% of the data was reserved for testing.
+        Figure~\ref{fig:solar_tft_overviwew} shows how the observed inputs, the irradiance measurements, along with corresponding data from historical irradiance forecasts are used as training inputs...
+
+    \subsubsection{Tuning}
+        Significant tuning was done to optimize the performance of both models.
+Since the goal of tuning was not only to minimize Mean Absolute Percentage Error, but balance that with the ability to recognize uncertainty, the process was very manual.
+Tables~\ref{tab:full_params} and \ref{tab:lim_params} show the selected hyperparameters for the full and limited models respectively.
 

### 2022-05-07 MSC-GUDMUNDSSON-2022.tex
Added to background and significant fixes throughout

@@
 %% online: for online versions (skip blank pages)
-\documentclass[online,masters,deptscs,forcegraphics,final]{ruthesis}
+\documentclass[online,masters,deptscs,forcegraphics,draft]{ruthesis}
 
@@
 Instability and unpredictability are undesirable in power production since production must match demand.
-In order to ensure power availability, highly polluting and expensive fossil power stations, \enquote{peaker plants}, are kept on standby to react to the instability in solar power.
+In order to ensure power availability when dependent on solar power, highly polluting and expensive fossil power stations, \enquote{peaker plants}, are kept on standby to react to the instability in solar power.
 The ability to reliably predict solar power production provides more predictability to power grid operators, allowing for greater efficiency in grid operation.
@@
 {
-Íslenska!
+Hraðar og hraðar erum við að færast frá því að framleiða raforku með jarðefnaeldsneyti, í átt að umhverfisvænni mátum til raforkuframleiðslu eins og sólarorku.
+Sólarorka er óútreiknanleg og mjög háð veðurfari, aðallega skýjahulu.
+Óstöðugleiki og óútreiknanleiki eru óákjósanlegir þegar kemur að sólarorku, þar sem framleiðsla verður að vera jöfn neyslu.
+Til að tryggja afhendingu raforku, þegar sólarorka er notuð, reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði.
+Áreiðanleg geta til að sjá fyrir um sólarorkuframleiðslu gerir stjórnun raforkukerfa einfaldari og útreiknanlegri, sem eykur skilvirkni þeirra.
+Verðmæt gögn eru til í formi veðurspáa, sem hægt er að nýta til að gera sólarorkuframleiðslu útreiknanlegri.
+Við notum nýjar aðferðir á sviði djúpnáms, athygli(e.
+attention), og þá sérstaklega tímasamrunabylta (e.
+temporal fusion transformers) til að spá fyrir um sólargeislun við sólarorkuver, sem fylgir náið raforkuframleiðslu orkuversins.
+Við byggjum þessa vinnu á háupplausnarveðurspám Belgings ehf., 24 klst.
+geislunarspám fyrir staðsetningu okuversins.
+Módelið nær meðal beinni meðalprósentuskekkju (e.
+mean absolute percentage error) upp á 13.21\% á prófunargagnasettinu fyrir eins dags spár.
+Að auki skilar módelið öryggisbili fyrir spárnar, sem veitir raforkuframleiðendum innsæi um það hvort þeir geti búist við óútreiknanlegri sólarorkuframleiðslu.
+Módelið skilar nákvæmari spám þegar það styðst við veðurspágögn af stærra svæði í nánd orkuversins en þegar það nýtir eingöngu gögn fyrir nærumhverfi orkuversins.
 } % Icelandic abstract goes here
@@
 PhD &Doctor of Philosophy\\
+AI &Artificial Intelligence\\
 ANN &Artificial Neural Network\\
@@
 \include{introduction}%%RUM: Introduction
+\include{background}
 \include{methods}%%RUM: "Methods"

### 2022-05-07 background.tex
Added to background and significant fixes throughout

--- 2022-04-08 introduction.tex
@@
-\chapter{Introduction\label{cha:introduction}}
+\chapter{Background\label{cha:background}}
+In this chapter we discuss the work our work builds on to give our work context.
+We cover the history of neural networks in machine learning, previous architectures that have been used to solve similar problems and the components of the solution we arrived at.
 
-Solar power production is extremely dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016}.
-A single cloud can seriously impact the production of a solar farm.
-Such unreliability in a power production system is highly undesirable since production must match demand \cite{lee_forecasting_2018}.
-When solar power production falls, the power demand it was serving must be fulfilled by other means.
-Fossil fueled power plants are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
-These plants often take hours to start, generally pollute more than other plants and are very expensive to run \cite{lee_forecasting_2018}.
-Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}.
-This notice enables planning for the dips in power production and making arrangements for the future.
-This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
-Recent advancements in deep learning have unlocked new tools that have been shown to be extremely effective in predicting the future behaviour of a system given enough relevant data.\\
+\section{Neural Networks}
+Artificial Neural Networks (ANNs), generally known as just Neural Networks (NNs) were first theorized in the 1940s \cite{mcculloch_logical_1943}.
+They are supposed to emulate the function of brains.
+They have \enquote{neurons} which, when they receive a signal, apply a function to the signal which they relay to the next neuron.
+Neural networks are built in layers.
+A layer receives a vector of input values, each neuron in the layer takes a weighted sum of the inputs and applies its activation function to the result.
+These values are then passed on to the next layer.
+Training, a neural network is the process of optimizing the weights in the layers of the network such that a specified input yields the desired output \cite{goodfellow_deep_2016}.\\
+Frank Rosenblatt Perceptron in 1950s.
+A single layer of neurons with the ability to classify images a few hundred pixels in size.
+Train neurons based on a dataset.
+His training algorithm was very limited and so inefficient that it did not allow for neural networks with more than one layer \cite{rosenblatt_perceptron_1958}.
+First theorized ability to recognize people and translate between languages.\\
+In 1974 Paul Werbos proposed backpropagation to optimize neural networks \cite{Werbos:74}, which became popular in 1986  with the work of Rumelhart et al..
+Backpropagation is an efficient training algorithm opening up the possibility of training neural networks deeper than one layer using gradient methods\cite{noauthor_backpropagation_2022}.
+Backpropagation enabled the creation of deeper neural networks, notably used to read handwritten digits with good accuracy \cite{lecun_backpropagation_1989}.\\
+Although backpropagation enabled neural networks with multiple layers, it did not enable significant depth.
+The most commonly used activation functions of the time, sigmoid and tanh, have derivatives close to zero, the gradient in the network values flatten at an exponential rate.
+This slows down the gradient descent used to optimize the networks, also at an exponential rate.
+In the 2010s the ReLu activation function \cite{nair_rectified_2010} coupled with improved optimization algorithms like ADAM and massive advances in computer hardware opened the door for Deep Learning, neural networks with many more layers.
+Today, deep neural networks have been used to great effect in applications like Google DeepMind's AlphaGO \cite{silver_mastering_2016}, the first AI to beat the world champion in the game of GO, and OpenAI's GPT3 model generates text human-like enough to pass the Turing test, to be confused with text written by a human \cite{vandeput_brief_nodate}.
 
-%% \ifdraft only shows the text in the first argument if you are in draft mode.
-%% These directions will disappear in other modes.
-%\ifdraft{State the objectives of the exercise. Ask yourself:
-%  \underline{Why} did I design/create the item? What did I aim to
-%  achieve? What is the problem I am trying to solve?  How is my
-%  solution interesting or novel?}{}
-
-\section{Background}
-
-\subsection{LSTMs and GRUs}
-\textcolor{blue}{Is there value in going deeper?}\\
+\section{LSTMs and GRUs\label{cha:RNN}}
 Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
-RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently.
+RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently\cite{lim_time_2021}.
 LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.\\
@@
 
-\subsection{Transformers}
+\section{Transformers\label{cha:transformer}}
 %Hard to train
@@
 
-Transformers have been used for a while in Natural Language Processing (NLP) and have more recently been used very effectively in forecasting.\cite{vaswani_attention_2017} They can more effectively learn and preserve local context located far apart in the input data.
+Transformers have been used for a while in Natural Language Processing (NLP) and have more recently been used very effectively in forecasting \cite{vaswani_attention_2017}.
+They can more effectively learn and preserve local context located far apart in the input data.
 As an example from NLP, the training data can contain multiple cases where time is being told, far apart.
 A transformer model would then more easily identify that the word “o’clock” is generally preceded by a number, whereas RNNs would be more likely to struggle with such distributed information.
-This can be applied to our work as similar weather patterns would reoccur days apart and cause a similar effect on measured irradiance at the power station.
+This can be applied to our work as similar weather patterns would reoccur days apart and cause a similar effect on measured irradiance at the power station \cite{lim_time_2021}.
 
+\afterpage{%
 \begin{figure}[ht!]
     \centering
-    \includegraphics[scale=0.7]{imgs/transformer.png}
+    \includegraphics[scale=0.95]{imgs/transformer.png}
     \caption{The transformer architecture.\cite{vaswani_attention_2017}
@@
 \end{figure}
+\clearpage
+}
 
-\subsubsection{Encoder/decoder}
+\subsection{Encoder/decoder}
 Transformers have an encoder/decoder structure.
@@
 
-\subsubsection{Positional encoding}
+\subsection{Positional encoding}
 Unlike RNNs, the transformer architecture has no recurrence and does inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
 
-\subsubsection{Self-attention} We previously discussed transformers' ability to preserve context across the long input sequences.
+\subsection{Self-attention} We previously discussed transformers' ability to preserve context across the long input sequences.
 That is thanks to the attention layer.
@@
 A single attention head can put all the focus on a single element.
-Multiple attention heads can be used to let the transformer model consider the context of more elements at the same time.\cite{rohrer_transformers_2021}
+Multiple attention heads can be used to let the transformer model consider the context of more elements at the same time \cite{rohrer_transformers_2021}.
 
-\subsection{Weather forecasts}
+\section{Weather forecasts\label{cha:WRF}}
 \textcolor{red}{TODO}
+\cite{gueymard_global_2008, rognvaldsson_numerical_2013, jimenez_wrf-solar_2016}
 
-\subsection{Previous work}
+\section{Previous work}
 Lin et al.
@@
 A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.\\
-
-\section{Goals}
-The methods previously used all have in common that they must effectively predict the weather so they can predict power output.
-
-Here we explore using high-resolution weather forecast data from meteorological models, offloading a majority of this difficult task to the field of meteorology.
-Instead, we focus on correlating these forecasts with measurements and detecting and reducing temporal and spatial error in the general forecasts using newer methods like Temporal Fusion Transformers.
-What is the quality of the predictions this methodology produces? Can we create a model which is useful to power producers in operating their systems more efficiently?
-
-%Here we explore whether better prediction accuracy can be reached
-%if high quality, high resolution weather forecast data is used. This offloads a majority of the difficult task of predicting the weather from the responsibilities of the network. \\
-%Can we construct a transformer based neural network which, by utilizing weather forecast data, is superior to neural networks which only utilize historical weather data?\\
-
-%% Glossary is broken, do not use --foley
-% \gls{auv}\footnote{Autonomous Undersea Vehicle}.
-
-% Notice that there is now information on the AUV in the Index and Acronyms.
-% It isn't in the \gls{glossary} because we didn't put it there.
-%\index{AUV}
-%}{}

### 2022-05-07 introduction.tex
Added to background and significant fixes throughout

--- 2022-04-25 introduction.tex
@@
 When solar power production falls, the power demand it was serving must be fulfilled by other means.
-Fossil fueled power plants are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
-These plants often take hours to start, generally pollute more than other plants and are expensive to run \cite{lee_forecasting_2018}.
+Fossil fueled power stations are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
+These power stations often take hours to start, generally pollute more than other power stations and are expensive to run \cite{lee_forecasting_2018}.
 Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}.
@@
 
-\section{Background}
-
-\textcolor{red}{TODO: Background on neural networks}\\
-
-\subsection{LSTMs and GRUs}
-Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
-RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently.
-LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.\\
-
-%\ifdraft{Provide background about the subject matter (e.g. How was morse code
-%developed?  How is it used today?). 
-%This is a place where there are usually many citations.
-%It is suspicious when there is not.
-%Include the purpose of the different equipment and your design intent. 
-%Include references to relevant scientific/technical work and books.
-%What other examples of similar designs exist?
-%How is your approach distinctive?
-
-%If you have specifications or related standards, these must be
-%described and cited also.  As an example, you might cite the specific
-%RoboSub competition website (and documents) if working on the lighting system for an AUV\cite{guls2016auvlight}
-
-\subsection{Transformers}
-%Hard to train
-%sequential
-% -> parallelization hard
-%encoder/decoder instead. Attention. 
-% Explain enc/dec
-% Autoencoders generate same output as the input, so no labeled data is needed -> completely unsupervisd earning.
-% Locality a problem with CNNs and RNNs. Can easily find local context but keeping on to context far apart in input is difficult. Enter transformers.
- %Encoder reads blocks of data and uses that and previously generated data as inputs into decoder which generatesdata one point at a time.
- %Attention. How you relate things to each other. i.e. input to output
- %self attention. How do features in the input data relate to each other
- \textcolor{red}{TODO: Make this coherent}
-
-Transformers have been used for a while in Natural Language Processing (NLP) and have more recently been used very effectively in forecasting.\cite{vaswani_attention_2017} They can more effectively learn and preserve local context located far apart in the input data.
-As an example from NLP, the training data can contain multiple cases where time is being told, far apart.
-A transformer model would then more easily identify that the word “o’clock” is generally preceded by a number, whereas RNNs would be more likely to struggle with such distributed information.
-This can be applied to our work as similar weather patterns would reoccur days apart and cause a similar effect on measured irradiance at the power station.
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[scale=0.7]{imgs/transformer.png}
-    \caption{The transformer architecture.\cite{vaswani_attention_2017}
-    \label{fig:transformer}}
-\end{figure}
-
-\subsubsection{Encoder/decoder}
-Transformers have an encoder/decoder structure.
-An encoder takes an input sequence and generates an abstract hidden state which the decoder then uses to generate an output sequence.
-A major advantage of this methodology is that the inputs and outputs do not have to have the same length, whereas the primary limit of the architecture is that all information in the input sequence needs to be represented by a single vector for the decoder to decode.
-The encoder distils the input down to a much smaller representation, a bottleneck, only containing desired information which the decoder in turn extracts.
-This is on its own used for tasks like de-noising or colourizing images.
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[scale=0.5]{imgs/encoder_decoder.jpeg}
-    \caption{The encoder/decoder structure
-    \label{fig:encoder_decoder}}
-\end{figure}
-
-\subsubsection{Positional encoding}
-Unlike RNNs, the transformer architecture has no recurrence and does inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
-
-\subsubsection{Self-attention} We previously discussed transformers' ability to preserve context across the long input sequences.
-That is thanks to the attention layer.
-The attention layer encodes an input data point's relevance to the rest of the points in the input sequence.
-Calling back to the previous NLP example, it encodes the context that the word "o'clock" is relevant to numbers because it is preceded by a number in most inputs.
-
-A single attention head can put all the focus on a single element.
-Multiple attention heads can be used to let the transformer model consider the context of more elements at the same time.\cite{rohrer_transformers_2021}
-
-\subsection{Weather forecasts}
-\textcolor{red}{TODO}
-
-\subsection{Previous work}
-Lin et al.
-\cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on the task, Temporal Convolutional Neural Networks (TCNN) can show even better performance.
-\textit{"TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
-and residual connections"} \cite{lin_temporal_2020}. \\
-Jaidee et al.
-\cite{jaidee_very_2019} showed that LSTMs, GRUs and derivative methods show relatively similar performance for very short-term predictions on the time scale of a few hours.
-This timescale is where it is critical to accurately predict cloud movements.
-These methods have been taken as far as building genetic algorithms to generate the optimal neural network for the purpose \cite{jaidee_very_2019}.
-These genetic algorithms operate on the same principles as animal breeding.
-They generate multiple candidate neural networks and use the best networks as a base to generate new networks for the next generation.
-This is done for multiple generations and in the end, the best neural network generated is used \cite{jaidee_very_2019}.\\
-Su et al.
-\cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto Regressive Neural Network (NARXNN), as well as various statistical approaches.
-When predicting power output, NARXNN takes in the current power output values and considers them as well as the past values of the power output of the system.
-These experiments showed that NARXNN had the best performance of these, by a rather large margin, and a hybrid method, of the better performing methods, showed even better results \cite{anderson_using_2018}.
-The statistical methods generally performed significantly worse in predicting the power output.
-\\
-Satellite imagery has been used for predicting solar power output as well.
-A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.\\
-
 \section{Goals}
-The methods previously used all have in common that they must effectively predict the weather so they can predict power output.
+Previous methods utilized in the prediction of solar power production all have in common that they must effectively predict the weather so they can predict power output.
 
@@
 The thesis is structured as follows.
-First we talked about the motivations and background in the introduction.
-Next we cover the methods, the data and tools used to achieve these goals.
+First we talked about the motivations for the work.
+Next we discuss the background the work stands on.
+Next we cover the methods utilized, the data and tools used to achieve these goals.
 The proceeding chapter contains the results of the work, and finally we discuss the implications, strengths and flaws of the work, future work and end with conclusions.

### 2022-05-07 methods.tex
Added to background and significant fixes throughout

@@
 \chapter{Methods}
+
+In this chapter we discuss our solution to the problem of predicting solar power production.
+We cover the data we had access to and ultimately selected to train the neural network with and how we processed and structured the data.
+We explain the Temporal Fusion Transformer architecture and how we set it up and use it to achieve our goals.
 
@@
 These irradiance values were normalized linearly to the range $0.0-1.0$ before input into the neural network.
-
-\subsubsection{Data Source}\label{cha:WRF}
-\textcolor{red}{TODO: Talk about WRF} \cite{gueymard_global_2008, rognvaldsson_numerical_2013, jimenez_wrf-solar_2016}
 
@@
 
+        \vspace{1.5cm}
+
         \begin{table}[ht!]

### 2022-05-07 results.tex
Added content to methods, added abstract, cleanup, new diagrams and more.
Added to background and significant fixes throughout

@@
     \centering
-    \includegraphics[scale=0.75]{imgs/graphs/full/days_full.png}
+    \includegraphics[scale=1]{imgs/graphs/full/days_full.png}
     \caption{The recorded Mean Absolute Percentage Error (MAPE) for each day in the test dataset for the full model. 
@@
 These predictions have a higher MAPE of 10-13\%.
+
+\subsection{Medium variance days}
+On days with some variance in the irradiance measured, the model is generally able to predict the irradiance with relatively good accuracy.
+The model tends to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value is within the 90\% confidence interval.
+
+Figure~\ref{fig:full_med_best} shows examples of good predictions with high confidence, though the confidence is generally lower the farther the irradiance is from the value of a clear day.
+
+\afterpage{%
 \begin{figure}[ht!]
@@
 
-\subsection{Medium variance days}
-On days with some variance in the irradiance measured, the model is generally able to predict the irradiance with relatively good accuracy.
-The model tends to give a smoother prediction than the measured value, averaging through the peaks and valleys of the measured values and most of the time the observed value is within the 90\% confidence interval.
-
-Figure~\ref{fig:full_med_best} shows examples of good predictions with high confidence, though the confidence is generally lower the farther the irradiance is from the value of a clear day.
 \begin{figure}[ht!]
@@
 \end{figure}
+\clearpage
+}
 
@@
 For Figures~\ref{fig:full_high_a}, \ref{fig:full_high_b} the sum of predicted values is close to the sum of observed values, indicating relatively high accuracy for the days taken as a whole.
-Figure\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that received very little irradiance.
+Figure~mi\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that received very little irradiance.
 \begin{figure}[ht!]
@@
     \centering
-    \includegraphics[scale=0.75]{imgs/graphs/less/days_single.png}
+    \includegraphics[scale=1]{imgs/graphs/less/days_single.png}
     \caption{The recorded Mean Absolute Percentage Error for each day in the test dataset of the limited model. 
@@
 
+\newpage
 \subsection{Low variance days}
@@
 \end{figure}
-poopies manana
 
@@
 Figure~\ref{fig:less_med_c} shows the opposite, the model predicts low irradiance for the whole day with a very large 90\% confidence interval of roughly $\pm200 W/m^2$.
+
+\afterpage{%
+\vspace{5cm}
 \begin{figure}[ht!]
     \centering
-    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}\label{fig:less_med_a}}\qquad
-    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_2.png}\label{fig:less_med_b}}\qquad
-    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_1.png}\label{fig:less_med_c}}\qquad
+    \makebox[\textwidth][c]{\subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}\label{fig:less_med_a}}\qquad}
+    \makebox[\textwidth][c]{\subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_2.png}\label{fig:less_med_b}}\qquad}
+    \makebox[\textwidth][c]{\subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_1.png}\label{fig:less_med_c}}\qquad
+    }
     \caption{Examples of days with some drops in irradiance.

### 2022-05-08 MSC-GUDMUNDSSON-2022.tex
General fleshing out, refinement and fixes.

@@
 Valuable data exists in weather forecasts, which can be harnessed to increase predictability in solar power generation.
-We use novel deep learning methods, attention, and specifically Temporal Fusion Transformers, to forecast the irradiance at a solar power station, which correlates very strongly with power output.
+We use novel deep learning methods, attention, and specifically temporal fusion transformers, to forecast the irradiance at a solar power station, which correlates very strongly with power output.
 We base this on Belgingur ehf.'s high-resolution weather forecast data, the 24-hour forecast irradiance values for the area of the power station.
-Our model achieves a very low average Mean Absolute Percentage Error of 13.21\% on the test dataset for one-day forecasts.
+Our model achieves a very low average mean absolute percentage error of 13.21\% on the test dataset for one-day forecasts.
 In addition, the model outputs a confidence level for its predictions providing insight to a grid operator on whether to expect unpredictability in the power output.
@@
 geislunarspám fyrir staðsetningu okuversins.
-Módelið nær meðal beinni meðalprósentuskekkju (e.
-mean absolute percentage error) upp á 13.21\% á prófunargagnasettinu fyrir eins dags spár.
+Módelið nær meðal meðaltölugildisprósentuskekkju (e.
+mean absolute percentage error) upp á 13,21\% á prófunargagnasettinu fyrir eins dags spár.
 Að auki skilar módelið öryggisbili fyrir spárnar, sem veitir raforkuframleiðendum innsæi um það hvort þeir geti búist við óútreiknanlegri sólarorkuframleiðslu.
@@
 \begin{dedications}
-  I dedicate this to my spouse/child/pet/power animal.
+  I dedicate this to my dog Birta. Rest in peace.
 \end{dedications}
@@
 
-\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
-  Note that it is important to acknowledge any funding that helped in the work}
+I would like to thank my supervisor, Yngvi, for invaluable advice throughout this work and Ólafur and Jorge at Belgingur for their help with understanding the weather and solar power aspect of the problem.
+I would also like to thank my friends, family and fellow thesis writers for being around and keeping me company while I was working on this.\\
 
@@
 \begin{tabular}{ll}
-MSc &Masters of Science\\
+MSc &Master of Science\\
 PhD &Doctor of Philosophy\\

### 2022-05-08 background.tex
General fleshing out, refinement and fixes.

@@
 \chapter{Background\label{cha:background}}
-In this chapter we discuss the work our work builds on to give our work context.
+In this chapter, we discuss the work our work builds on to give our work context.
 We cover the history of neural networks in machine learning, previous architectures that have been used to solve similar problems and the components of the solution we arrived at.
@@
 Artificial Neural Networks (ANNs), generally known as just Neural Networks (NNs) were first theorized in the 1940s \cite{mcculloch_logical_1943}.
-They are supposed to emulate the function of brains.
-They have \enquote{neurons} which, when they receive a signal, apply a function to the signal which they relay to the next neuron.
+They are supposed to emulate the function of biological brains.
+They are composed of \enquote{neurons} which, when they receive a signal, apply a function to the signal which they then relay to the next neuron.
 Neural networks are built in layers.
-A layer receives a vector of input values, each neuron in the layer takes a weighted sum of the inputs and applies its activation function to the result.
+A layer receives a tensor of input values, each neuron in the layer takes a weighted sum of the inputs and applies its activation function to the result.
 These values are then passed on to the next layer.
-Training, a neural network is the process of optimizing the weights in the layers of the network such that a specified input yields the desired output \cite{goodfellow_deep_2016}.\\
-Frank Rosenblatt Perceptron in 1950s.
-A single layer of neurons with the ability to classify images a few hundred pixels in size.
-Train neurons based on a dataset.
+Training a neural network is the process of optimizing the weights in the layers of the network such that a specified input yields the desired output via iteration to minimize the output of a loss function \cite{goodfellow_deep_2016}.\\
+Frank Rosenblatt created the perceptron in the 1950s.
+A perceptron is a single layer of neurons with the ability to classify images a few hundred pixels in size.
+Rosenblatt's breakthrough idea was to train a neural network based on reinforcement from a dataset, strengthening successful connections and weakening unsuccessful connections.
 His training algorithm was very limited and so inefficient that it did not allow for neural networks with more than one layer \cite{rosenblatt_perceptron_1958}.
-First theorized ability to recognize people and translate between languages.\\
-In 1974 Paul Werbos proposed backpropagation to optimize neural networks \cite{Werbos:74}, which became popular in 1986  with the work of Rumelhart et al..
-Backpropagation is an efficient training algorithm opening up the possibility of training neural networks deeper than one layer using gradient methods\cite{noauthor_backpropagation_2022}.
+Based on his work, Rosenblatt first theorized the ability of neural networks to recognize people and translate between languages \cite{vandeput_brief_nodate}.\\
+In 1974 Paul Werbos proposed backpropagation to optimize neural networks \cite{Werbos:74}, which became popular in 1986  with the work of Rumelhart et al.
+\cite{rumelhart_learning_1986}.
+Backpropagation is an efficient training algorithm opening up the possibility of training neural networks deeper than one layer using gradient descent.
+\enquote{The backpropagation algorithm works by computing the gradient of the loss function with respect to each weight by the chain rule, computing the gradient one layer at a time, iterating backward from the last layer to avoid redundant calculations of intermediate terms in the chain rule} \cite{noauthor_backpropagation_2022}.
 Backpropagation enabled the creation of deeper neural networks, notably used to read handwritten digits with good accuracy \cite{lecun_backpropagation_1989}.\\
 Although backpropagation enabled neural networks with multiple layers, it did not enable significant depth.
-The most commonly used activation functions of the time, sigmoid and tanh, have derivatives close to zero, the gradient in the network values flatten at an exponential rate.
+The most commonly used activation functions of the time, sigmoid and tanh, have derivatives close to zero, so stacking them flattens the gradient in the network values at an exponential rate.
 This slows down the gradient descent used to optimize the networks, also at an exponential rate.
 In the 2010s the ReLu activation function \cite{nair_rectified_2010} coupled with improved optimization algorithms like ADAM and massive advances in computer hardware opened the door for Deep Learning, neural networks with many more layers.
-Today, deep neural networks have been used to great effect in applications like Google DeepMind's AlphaGO \cite{silver_mastering_2016}, the first AI to beat the world champion in the game of GO, and OpenAI's GPT3 model generates text human-like enough to pass the Turing test, to be confused with text written by a human \cite{vandeput_brief_nodate}.
+Today, deep neural networks have been used to great effect in applications like Google DeepMind's AlphaGO \cite{silver_mastering_2016}, the first AI to beat the world champion in the game of go, and OpenAI's GPT3 model generates text human-like enough to pass the Turing test, to be confused with text written by a human \cite{vandeput_brief_nodate}.
 
-\section{LSTMs and GRUs\label{cha:RNN}}
-Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long-Short Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
-RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently\cite{lim_time_2021}.
-LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.\\
+\section{Recurrent Neural Networks\label{cha:RNN}}
+Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long Short-Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016, hochreiter_long_1997} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
+RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently \cite{lim_time_2021}.
+LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.
+Figure~\ref{fig:RNNs} shows the structure of the nodes of the three different Recurrent Neural Network types discussed above.\\
+
+\begin{figure}[ht!]
+    \centering
+    \subfloat[Plain Recurrent Neural Network (RNN)]{\includegraphics[scale=0.35]{imgs/RNN.png}\label{fig:RNN}}\qquad
+    \subfloat[Gated Recurrent Unit (GRU)]{\includegraphics[scale=0.35]{imgs/GRU.png}\label{fig:GRU}}\qquad
+    \subfloat[Long Short-Term Memory (LSTM)]{\includegraphics[scale=0.35]{imgs/LSTM.png}\label{fig:LSTM}}\qquad
+    \caption{The cells of different Recurrent Neural Network types.
+$x_t$ represents inputs, $h_t$ and $C_t$ represent hidden layer vectors passed between cells.
+$\sigma$ and $tanh$ are activation functions.
+\cite{lopez_rnn_nodate}.
+    \label{fig:RNNs}}
+\end{figure}
 
@@
  %self attention. How do features in the input data relate to each other
- \textcolor{red}{TODO: Make this coherent}
 
@@
 As an example from NLP, the training data can contain multiple cases where time is being told, far apart.
-A transformer model would then more easily identify that the word “o’clock” is generally preceded by a number, whereas RNNs would be more likely to struggle with such distributed information.
+A transformer model would then more easily identify that the word \enquote{o’clock} is generally preceded by a number, whereas RNNs would be more likely to struggle with such distributed information.
 This can be applied to our work as similar weather patterns would reoccur days apart and cause a similar effect on measured irradiance at the power station \cite{lim_time_2021}.
+Figure~\ref{fig:transformer} shows an overview of the transformer architecture.
 
 \afterpage{%
-\begin{figure}[ht!]
+\begin{figure}
     \centering
-    \includegraphics[scale=0.95]{imgs/transformer.png}
+    \includegraphics[scale=0.80]{imgs/transformer.png}
     \caption{The transformer architecture.\cite{vaswani_attention_2017}
@@
 \subsection{Encoder/decoder}
-Transformers have an encoder/decoder structure.
+Transformers have an encoder/decoder structure, depicted in Figure~\ref{fig:encoder_decoder}.
 An encoder takes an input sequence and generates an abstract hidden state which the decoder then uses to generate an output sequence.
@@
 
-\begin{figure}[ht!]
+\begin{figure}
     \centering
@@
 \subsection{Positional encoding}
-Unlike RNNs, the transformer architecture has no recurrence and does inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
+Unlike RNNs, the transformer architecture has no recurrence, or cell based memory, and does not inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
 
@@
 Su et al.
-\cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto Regressive Neural Network (NARXNN), as well as various statistical approaches.
-When predicting power output, NARXNN takes in the current power output values and considers them as well as the past values of the power output of the system.
+\cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto Regressive Neural Networks (NARXNN), as well as various statistical approaches.
+When predicting power output, NARXNN takes in the current power output values and considers them, in real time, as well as the past values of the power output of the system.
 These experiments showed that NARXNN had the best performance of these, by a rather large margin, and a hybrid method, of the better performing methods, showed even better results \cite{anderson_using_2018}.
@@
 A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.\\
+\textcolor{red}{AlexNet}

### 2022-05-08 discussion.tex
Implemented feedback from instructor
General fleshing out, refinement and fixes.

@@
 
-\section{Conclusion\label{sec:conclusions}}
-The results clearly show that relying on forecast irradiance data from meteorological models is a very effective, but not perfect, method of accurately predicting irradiance in a single area.
-A transformer model, the Temporal Fusion Transformer, managed to very effectively translate patterns in meteorological irradiance forecasts to accurate local irradiance predictions utilizing spatially and temporally adjacent data.
-
+\section{Limitations and Future Work}
 We used a dataset of a fairly limited size, only 425 days.
@@
 
+The model also has huge potential in utilizing more varied data in addition to a larger time span, other exogenous data that might further help relate the input forecast to the actual irradiance.
+Experimenting with wind data, both historical and forecast data, may well unlock further steps in performance.
+
 Sadly, there is not a standard benchmark for solar irradiance forecast models, and the data used by others who have done similar work is not available, so we can't make a proper comparison between methodologies and models.
+
+\section{Conclusion\label{sec:conclusions}}
+The results clearly show that relying on forecast irradiance data from meteorological models is a very effective, but not perfect, method of accurately predicting irradiance in a single area.
+A transformer model, the Temporal Fusion Transformer, managed to very effectively translate patterns in meteorological irradiance forecasts to accurate local irradiance predictions utilizing spatially and temporally adjacent data.\\
 %%% Local Variables: 

### 2022-05-08 introduction.tex
General fleshing out, refinement and fixes.

@@
 \chapter{Introduction\label{cha:introduction}}
+
+\vspace{1.5cm}
 
@@
 This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
-Recent advancements in deep learning have unlocked new tools that have been shown to be effective in predicting the future behaviour of a system given enough relevant data.\\
+Recent advancements in deep learning have unlocked new tools for forecasting, such as attention mechanisms, that have been shown to be effective in predicting the future behaviour of a system, given enough relevant data.\\
 
@@
 
+ \newpage
 \section{Goals}
@@
 Here we explore using high-resolution weather forecast data from meteorological models, offloading a majority of this difficult task to the field of meteorology.
-Instead, we focus on correlating these forecasts with measurements and detecting and reducing temporal and spatial error in the general forecasts using newer methods like Temporal Fusion Transformers.
+Instead of generating forecasts, we focus on correlating existing forecasts with measurements and detecting and reducing temporal and spatial error in the general forecasts using newer methods like Temporal Fusion Transformers.
 What is the quality of the predictions this methodology produces? Can we create a model which is useful to power producers in operating their systems more efficiently?
 
-The thesis is structured as follows.
-First we talked about the motivations for the work.
-Next we discuss the background the work stands on.
-Next we cover the methods utilized, the data and tools used to achieve these goals.
-The proceeding chapter contains the results of the work, and finally we discuss the implications, strengths and flaws of the work, future work and end with conclusions.
+This thesis is structured as follows.
+First, we talked about the motivations for the work.
+Next, we discuss the background the work stands on.
+After that, we cover the methods utilized, the data and tools used to achieve these goals.
+The proceeding chapter contains the results of the work, and finally, we discuss the implications, strengths and flaws of the work, future work and end with conclusions.
 

### 2022-05-08 methods.tex
General fleshing out, refinement and fixes.

@@
 
-In this chapter we discuss our solution to the problem of predicting solar power production.
-We cover the data we had access to and ultimately selected to train the neural network with and how we processed and structured the data.
+In this chapter, we discuss our solution to the problem of predicting solar power production.
+We cover the data we had access to and ultimately selected to train the neural network with, and how we processed and structured the data.
 We explain the Temporal Fusion Transformer architecture and how we set it up and use it to achieve our goals.
@@
 This data is provided with five-minute intervals, but the neural network operates hourly.
-To downsample the measurement data to an hourly format, we averaged the values for the twenty minutes before and after the top of each hour.
-Finally, the irradiance values were normalized linearly to the range $0.0-1.0$ before input into the neural network.
+To downsample the measurement data to an hourly format, we averaged the values for the twenty minutes before and after the beginning of each hour.
+Finally, the irradiance values were normalized linearly to the range $0.0-1.0$ before being input into the neural network.
 
 \subsection{Forecast Data\label{cha:forecast_data}}
-The weather forecast data is provided by Belgingur ehf., Reykjavik, Iceland.
+The weather forecast data is provided by Belgingur ehf., Reykjavík, Iceland.
 We use irradiance values from historical 24-hour weather forecasts generated by their WRF model, discussed in Section~\ref{cha:WRF}.
@@
 We use two irradiance values from the forecast, the downward irradiance and the downward diffuse irradiance.
-These irradiance values were normalized linearly to the range $0.0-1.0$ before input into the neural network.
+These irradiance values were normalized linearly to the range $0.0-1.0$ before being input into the neural network.
 
@@
 Hence we do not evaluate early morning or late evening performance.
-While a valuable metric, MAPE does not  fully represent the value of the model.
+While a valuable metric, MAPE does not  fully represent the function of the model.
 Other factors like confidence, discussed in Section~\ref{sec:loss_function} and deviance from total irradiance for the day also represent key factors in the usefulness of the model.
@@
     \subsection{Temporal Fusion Transformer}
-    \textcolor{red}{TODO: more}
+    \textcolor{red}{TODO: add more meat}
 
@@
     \subsection{Setup}
-    \textcolor{red}{TODO: finish --- early stopping}\\
     We trained two models.
-One which utilizes weather forecast data from an area surrounding the power station, and one which does not, discussed in Section~\ref{cha:forecast_data}.
-The limited model is intended to show that the model is able to glean additional information from the surrounding area data.
+One which utilizes weather forecast data from an area surrounding the power station, and one which does not, as discussed in Section~\ref{cha:forecast_data}.
+The limited model is intended to show that the full model is able to glean additional information from the surrounding area data.
 
@@
     \subsubsection{Inputs}
-        The dataset was split into a training set, composing 70\% of the data, a validation set, composing 10\% of the data and finally 20\% of the data was reserved for testing.
-        Figure~\ref{fig:solar_tft_overviwew} shows how the observed inputs, the irradiance measurements, along with corresponding data from historical irradiance forecasts are used as training inputs...
+        The dataset was split into a training set, composing 70\% of the data, a validation set, composing 10\% of the data and finally, 20\% of the data was reserved for testing.
+        Figure~\ref{fig:solar_tft_overviwew} shows how the observed inputs, the irradiance measurements, along with corresponding data from historical irradiance forecasts are used as training inputs.
 
@@
         Significant tuning was done to optimize the performance of both models.
-Since the goal of tuning was not only to minimize Mean Absolute Percentage Error, but balance that with the ability to recognize uncertainty, the process was very manual.
-Tables~\ref{tab:full_params} and \ref{tab:lim_params} show the selected hyperparameters for the full and limited models respectively.
+Since the goal of tuning was not only to minimize Mean Absolute Percentage Error but balance that with the ability to recognize uncertainty, the process was very manual.
+Tables~\ref{tab:full_params} and \ref{tab:lim_params} show the selected hyperparameters for the full and limited models respectively.\\
+
+        To avoid overfitting and to speed up the training of models, we used early stopping to stop training when the loss function optimization plateaued.
 
@@
         Dropout                 & 0.1            \\ \hline
-        Hidden size            & 64         \\ \hline
+        Gradient clip           & 0.1           \\ \hline
+        Hidden layer size            & 64         \\ \hline
         Hidden continuous size & 8         \\ \hline
@@
         Dropout                 & 0.1            \\ \hline
-        Hidden size            & 40         \\ \hline
+        Gradient clip           & 0.1           \\ \hline
+        Hidden layer size            & 40         \\ \hline
         Hidden continuous size & 6         \\ \hline
@@
     \subsubsection{Loss function}\label{sec:loss_function}
-    \textcolor{red}{TODO: Fix, find more sources and make sure this is correct}\\
+    \textcolor{red}{TODO: Flesh out and fix}\\
     We use the quantile loss  Temporal Fusion Transformers is quantile loss, which we use.

### 2022-05-08 results.tex
General fleshing out, refinement and fixes.

@@
 15.98\%, as well as better tracking in challenging conditions and a better ability to recognize uncertain conditions, generally reporting lower confidence when its predictions are wrong.
-Both models can be trained to achieve slightly lower MAPE, but that comes at the cost of a higher tendency to underfit variance, and higher false confidence.
+Both models can be trained to achieve slightly lower MAPE, but that comes at the cost of a higher tendency to underfit variance and higher false confidence.
 We show irradiance plots for select days representative of the models' performance in various conditions.
@@
     \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/full/low/best_2.png}}\qquad
-    \caption{Examples of clear days where the model preforms very well.
+    \caption{Examples of clear days where the model performs very well.
     \label{fig:full_low_best}}
@@
 The model does not manage near-perfect predictions on all of these clear days, and makes a few under-predictions, as can be seen in Figure~\ref{fig:full_low_worst}.
-These predictions still predict steady sunshine, and are within the 90\% confidence interval, which is slightly larger than for the other clear day predictions.
+These predictions still predict steady sunshine and are within the 90\% confidence interval, which is slightly larger than for the other clear day predictions.
 These predictions have a higher MAPE of 10-13\%.
@@
 \afterpage{%
-\begin{figure}[ht!]
+\begin{figure}
     \centering
@@
 For Figures~\ref{fig:full_high_a}, \ref{fig:full_high_b} the sum of predicted values is close to the sum of observed values, indicating relatively high accuracy for the days taken as a whole.
-Figure~mi\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that received very little irradiance.
+Figure~\ref{fig:full_high_d} shows an outlier with an extremely high MAPE of 139\%, where the model predicts an almost clear day, with high confidence, on a day that received very little irradiance.
 \begin{figure}[ht!]
@@
 \section{Less forecast data}
-To confirm that the full model utilizes the input data for the surrounding area, we also trained the model on only irradiance forecast for the cell the power station is located in.
+To confirm that the full model utilizes the input data for the surrounding area, we also trained the model on only the irradiance forecast for the cell the power station is located in.
 The limited model achieves an average MAPE of 15.98\% on the test dataset.
@@
 \vspace{5cm}
-\begin{figure}[ht!]
+\begin{figure}
     \centering
@@
 \subsection{High variance days}
-For the few days in the dataset that had very low irradiance for part of, or the whole day or large fluctuations in irradiance, the model displays high error, MAPE in the range 53-123\% with varying confidence.
+For the few days in the dataset that had very low irradiance for part of, or the whole day, or large fluctuations in irradiance, the model displays a high error, MAPE in the range 53-123\% with varying confidence.
 Figures~\ref{fig:less_high_a} and \ref{fig:less_high_b} show large over-predictions with high confidence, and Figures~\ref{fig:less_high_c} and \ref{fig:less_high_d} show days where the 90\% confidence interval covers the top 60\% of the graph, yet it does not cover the observations.

### 2022-05-09 MSC-GUDMUNDSSON-2022.tex
Added weather forecast background. More cleanup.

@@
 
-I would like to thank my supervisor, Yngvi, for invaluable advice throughout this work and Ólafur and Jorge at Belgingur for their help with understanding the weather and solar power aspect of the problem.
-I would also like to thank my friends, family and fellow thesis writers for being around and keeping me company while I was working on this.\\
+I would like to thank my supervisor, Yngvi, for invaluable advice throughout this work and Ólafur and Jorge at Belgingur for their help with understanding the nature of the solar weather data.
+I would also like to thank my friends, family and fellow thesis writers for their company and moral support while I was working on this thesis.\\
 
@@
 GRN &Gated Residual Network\\
+NWP &Numerical Weather Prediction\\
+WRF &Weather Research and Forecasting\\
+NCAR &National Center for Atmospheric Research\\
+NOAA &National Oceanic and Atmospheric Administration\\
+AFWA &Air Force Weather Agency\\
+NRL &Naval Research Laboratory\\
+OU &University of Oklahoma\\
+FAA &Federal Aviation Administration\\
 MAPE &Mean Absolute Percentage Error\\

### 2022-05-09 background.tex
Added weather forecast background. More cleanup.

@@
 Transformers have an encoder/decoder structure, depicted in Figure~\ref{fig:encoder_decoder}.
-An encoder takes an input sequence and generates an abstract hidden state which the decoder then uses to generate an output sequence.
+An encoder takes an input sequence and generates an abstract hidden state which the decoder then uses to generate an output sequence \cite{vaswani_attention_2017}.
 A major advantage of this methodology is that the inputs and outputs do not have to have the same length, whereas the primary limit of the architecture is that all information in the input sequence needs to be represented by a single vector for the decoder to decode.
 The encoder distils the input down to a much smaller representation, a bottleneck, only containing desired information which the decoder in turn extracts.
-This is on its own used for tasks like de-noising or colourizing images.
+This is on its own used for tasks like de-noising or colourizing images \cite{nechu_what_2020}.
 
@@
 The attention layer encodes an input data point's relevance to the rest of the points in the input sequence.
-Calling back to the previous NLP example, it encodes the context that the word "o'clock" is relevant to numbers because it is preceded by a number in most inputs.
+Calling back to the previous NLP example, it encodes the context that the word \enquote{o'clock} is relevant to numbers because it is preceded by a number in most inputs.
 
@@
 \section{Weather forecasts\label{cha:WRF}}
-\textcolor{red}{TODO}
-\cite{gueymard_global_2008, rognvaldsson_numerical_2013, jimenez_wrf-solar_2016}
-
+Numerical Weather Prediction (NWP) has used mathematical models of the oceans and atmosphere since the 1920s to predict the weather based on current conditions, although realistic results weren't achieved until the advent of computer simulations in the 1950s \cite{noauthor_numerical_2022}.
+The Weather Research and Forecasting (WRF) model is an NWP system created by the National Center for Atmospheric Research (NCAR), the National Oceanic and Atmospheric Administration (NOAA), the Air Force Weather Agency (AFWA), the Naval Research Laboratory (NRL), the University of Oklahoma (OU), and the Federal Aviation Administration (FAA) in the United States for both research and operational weather forecasting \cite{noauthor_weather_2022}.\\
+WRF-SOLAR is an NWP model based on WRF specifically designed to model values useful for solar power, including high frequency irradiance calculations, more accurate solar position algorithms, and more robust aerosol and particle simulations with regard to radiation \cite{jimenez_wrf-solar_2016}.
 \section{Previous work}
@@
 \cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on the task, Temporal Convolutional Neural Networks (TCNN) can show even better performance.
-\textit{"TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
-and residual connections"} \cite{lin_temporal_2020}. \\
+\textit{\endquote{TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
+and residual connections}} \cite{lin_temporal_2020}. \\
 Jaidee et al.

### 2022-05-09 introduction.tex
Added weather forecast background. More cleanup.

@@
 Fossil fueled power stations are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
-These power stations often take hours to start, generally pollute more than other power stations and are expensive to run \cite{lee_forecasting_2018}.
+These power stations often take hours to start, generally pollute more than other power stations and are expensive to run \cite{lee_forecasting_2018}.\\
 Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}.
 This notice enables planning for the dips in power production and making arrangements for the future.
-This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.
+This provides the tools to reduce the cost and pollution of power generation as well as increase the reliability of power delivery to consumers \cite{lee_forecasting_2018}.\\
+
 Recent advancements in deep learning have unlocked new tools for forecasting, such as attention mechanisms, that have been shown to be effective in predicting the future behaviour of a system, given enough relevant data.\\
@@
 Instead of generating forecasts, we focus on correlating existing forecasts with measurements and detecting and reducing temporal and spatial error in the general forecasts using newer methods like Temporal Fusion Transformers.
-What is the quality of the predictions this methodology produces? Can we create a model which is useful to power producers in operating their systems more efficiently?
+What is the quality of the predictions this methodology produces? Can we create a model which is useful to power producers in operating their systems more efficiently?\\
 

### 2022-05-09 methods.tex
Added weather forecast background. More cleanup.

@@
 \subsection{Irradiance measurements}
+Under certain circumstances irradiance values simulated after the fact can be accurate, however this does not apply in all conditions, and real measured values are much more reliable \cite{gueymard_global_2008}.
 The irradiance was measured with seven Kipp \& Zonen CM21 pyranometers placed around the power station.
@@
 
-    We utilize a Temporal Fusion Transformer (TFT) model for the irradiance prediction.
+    \noindent We utilize a Temporal Fusion Transformer (TFT) model for the irradiance prediction.
     TFTs are  general networks which use attention at their core to generate multi-horizon forecasts with the ability to consider
@@
     }
-    TFTs are composed of various components, each primarily addressing a single input type (i.e. static, known or observed inputs), aimed at maximizing the model versatility across various problems.
 
-    \textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data.\\
+    \noindent TFTs are composed of various components, each primarily addressing a single input type (i.e.
+static, known or observed inputs), aimed at maximizing the model versatility across various problems.
+
+    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data.\\
     A Gated Residual Network (GRN), shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}.\\
-    \indent\textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.\\
-    \indent\textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
-    \indent\textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
-    \indent\textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, show a confidence interval for the predictions \cite{lim_temporal_2020}. \\
+    \textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.\\
+    \textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
+    \textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
+    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, show a confidence interval for the predictions \cite{lim_temporal_2020}. \\
     Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.
@@
         \centering
-         \makebox[\textwidth][c]{\includegraphics[scale=0.37]{imgs/TFT.png}}
+         \makebox[\textwidth][c]{\includegraphics[scale=0.36]{imgs/TFT.png}}
         \caption{A detailed view of the architecture of the Temporal Fusion transformer.\cite{lim_temporal_2020}
@@
     We use the quantile loss  Temporal Fusion Transformers is quantile loss, which we use.
-    A model's prediction is a random value within a probability distribution \cite{}.
+    A model's prediction is a random value within a probability distribution \cite{wen_multi-horizon_2018}.
 If we make multiple predictions we can estimate this probability distribution by seeing within what range a selected proportion of the predictions fall \cite{koenker_quantile_2001}.
 PyTorch Forecasting's quantile loss function takes the Mean Absolute Error (MAE) of three quantiles, 90\%, 50\% and 10\%.
-[0.02, 0.1, 0.25, 0.5, 0.75, 0.9, 0.98] This gives a good picture of the model's confidence in a prediction.
+This gives a good picture of the model's confidence in a prediction.
 A smaller quantile range indicated more confidence.

### 2022-05-09 discussion.tex
Complete content-wise. Waiting on final feedback.

@@
 The true measure of performance is how useful the model output is to energy producers.
-The model's confidence in a prediction is almost as important as the accuracy of the prediction and the total energy produced over the day matters.
+The model's confidence in a prediction is almost as important as the accuracy of the prediction, and the total energy produced over the day matters.
 The forecast downward shortwave irradiance, the largest factor in measured irradiance, rarely drops.
@@
 
-Fig.~\ref{fig:disc_low} shows side-by-side the two models' performance for the same calm day.
+Figure~\ref{fig:disc_low} shows side-by-side the two models' performance for the same calm day.
 The full model was able to track the observations more closely with more confidence.
-This was typical for all calm days.
+This was typical for all calm days.\\
+
 \begin{figure}[ht!]
@@
 
-In Fig.~\ref{fig:disc_med} we can see a comparison between the two models on a less predictable day.
+In Figure~\ref{fig:disc_med} we can see a comparison between the two models on a less predictable day.
 Again, the full model shows better performance in tracking, primarily because the limited model misses an end-of-day prediction.
@@
 
-Fig.~\ref{fig:disc_high} shows a clear example of the performance difference between the models on erratic days.
+Figure~\ref{fig:disc_high} shows a clear example of the performance difference between the models on erratic days.
 The full model, while still having a high error gets somewhat close to keeping the observations within its much smaller 90\% confidence interval than the limited model.

### 2022-05-09 introduction.tex
Complete content-wise. Waiting on final feedback.

@@
 
-Solar power production is highly dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016}.
-A single cloud can adversely impact the production of a solar farm.
+Solar power production is highly dependent on weather conditions \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019, jang_solar_2016, jimenez_wrf-solar_2016, rognvaldsson_numerical_2013}.
+A single cloud can adversely impact the production of a solar power station.
 Such unreliability in a power production system is undesirable since production must match demand \cite{lee_forecasting_2018}.

### 2022-05-09 methods.tex
Complete content-wise. Waiting on final feedback.

@@
     \subsection{Temporal Fusion Transformer}
-    \textcolor{red}{TODO: add more meat}
 
@@
 
-    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data.\\
-    A Gated Residual Network (GRN), shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}.\\
-    \textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.\\
+    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data, or datasets where it is unknown which parts are of significance.\\
+    The Gated Residual Network (GRN) structure, shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}.
+It provides the model the flexibility to choose the complexity of its input, more complex where needed, and less complex where not.\\
+    \textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.
+Some variables may be more important than others, \enquote{TFT is designed to provide instance-wise variable selection through the use of variable selection networks applied to both static covariates and time-dependent covariates.} \cite{lim_temporal_2020} The variable selection mechanism also allows the model to remove noisy inputs that might have a negative effect on performance.
+\\
     \textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
     \textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
-    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, show a confidence interval for the predictions \cite{lim_temporal_2020}. \\
+    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, display a confidence interval for the predictions \cite{lim_temporal_2020}. \\
     Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.
@@
     \subsubsection{Loss function}\label{sec:loss_function}
-    \textcolor{red}{TODO: Flesh out and fix}\\
-    We use the quantile loss  Temporal Fusion Transformers is quantile loss, which we use.
+    The loss function used for Temporal Fusion Transformers is quantile loss, as it's an integral part of generating the confidence intervals of the model \cite{lim_temporal_2020}.
     A model's prediction is a random value within a probability distribution \cite{wen_multi-horizon_2018}.
@@
 This gives a good picture of the model's confidence in a prediction.
-A smaller quantile range indicated more confidence.
+A smaller quantile range indicates more confidence.

### 2022-05-13 MSC-GUDMUNDSSON-2022.tex
Complete content-wise. Waiting on final feedback.
Small fixes

@@
 \supervisors{
-  \personinfo{Dr. Yngvi Björnsson, PhD}{\MLSupervisor}{\MLProfessor}{\MLReykjavikUniversity}{\MLIceland}
-  \personinfo{Dr. Ólafur Rögnvaldsson, PhD}{Examiner}{Managing Director}{Belgingur ehf.}{\MLIceland}
+  \personinfo{Yngvi
+  Björnsson}{\MLSupervisor}{\MLProfessor}{\MLReykjavikUniversity}{\MLIceland}
+
 %  \personinfo{Ian M. Great}{Co-advisor}{\MLProfessor}{Hochschule Düsseldorf}{Germany}
@@
 \examiners{
-  \personinfo{Tough E. Questions}{\MLExaminer}{Associate Professor}{Massachusetts Institute of Technology}{USA}
+  \personinfo{Ólafur Rögnvaldsson}{\MLExaminer}{Managing Director}{Belgingur ehf.}{\MLIceland}
+  \personinfo{Stephan Schiffel}{\MLExaminer}{Assistant Professor}{\MLReykjavikUniversity}{\MLIceland}
 
@@
 
-\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
-  have not already done so, \LaTeX{} the \path{instructions.tex} to
-  learn how to setup your document and use some of the features.  You
-  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
-  There is also more information on working with \LaTeX{} at
-  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
-  This includes common problems and fixes.
-
-  This page will disappear in anything other than draft mode.}{}
-
 %% Dedication is optional, comment out if it is absent
@@
 The project is partially funded by the Icelandic Technology Development Fund through grant nr. 1910034-0611.
-
-\coverchapter{Preface}
-% RUM: "Preface (optional)"
-This dissertation is original work by the author, Einar~Helgi~Guðmundsson.
-
-\draftnote{The preface is an optional element
-  explaining a little who performed what work.  See
-  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
-  for suggestions.
-
-  List of publications as part of the preface is
-  optional unless elements of the work have already been published.
-  It should be a comprehensive list of all publications in which
-  material in the thesis has appeared, preferably with references to
-  sections as appropriate.  This is also a good place to state
-  contribution of student and contribution of others to the work
-  represented in the thesis.}
 

### 2022-05-13 background.tex
Complete content-wise. Waiting on final feedback.
Small fixes

@@
 \section{Transformers\label{cha:transformer}}
+
+\afterpage{%
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.80]{imgs/transformer.png}
+    \caption{The transformer architecture.\cite{vaswani_attention_2017}
+    \label{fig:transformer}}
+\end{figure}
+\clearpage
+}
+
 %Hard to train
@@
 
-\afterpage{%
-\begin{figure}
-    \centering
-    \includegraphics[scale=0.80]{imgs/transformer.png}
-    \caption{The transformer architecture.\cite{vaswani_attention_2017}
-    \label{fig:transformer}}
-\end{figure}
-\clearpage
-}
-
 \subsection{Encoder/decoder}
@@
 
-\begin{figure}
+\begin{figure}[ht!]
     \centering
@@
 Lin et al.
-\cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on the task, Temporal Convolutional Neural Networks (TCNN) can show even better performance.
+\cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on solar power prediction, Temporal Convolutional Neural Networks (TCNN) can show even better performance.
 \textit{\endquote{TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
@@
 Jaidee et al.
-\cite{jaidee_very_2019} showed that LSTMs, GRUs and derivative methods show relatively similar performance for very short-term predictions on the time scale of a few hours.
+\cite{jaidee_very_2019} showed that LSTMs, GRUs and derivative methods show relatively similar performance for very short-term predictions of solar power production on the time scale of a few hours.
 This timescale is where it is critical to accurately predict cloud movements.
@@
 \cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto Regressive Neural Networks (NARXNN), as well as various statistical approaches.
-When predicting power output, NARXNN takes in the current power output values and considers them, in real time, as well as the past values of the power output of the system.
+When predicting solar power output, NARXNN takes in the current power output values and considers them, in real time, as well as the past values of the power output of the system.
 These experiments showed that NARXNN had the best performance of these, by a rather large margin, and a hybrid method, of the better performing methods, showed even better results \cite{anderson_using_2018}.
@@
 A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.\\
-\textcolor{red}{AlexNet}

### 2022-05-13 methods.tex
Small fixes

@@
 It provides the model the flexibility to choose the complexity of its input, more complex where needed, and less complex where not.\\
-    \textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.
+    \clearpage
+    \noindent\textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.
 Some variables may be more important than others, \enquote{TFT is designed to provide instance-wise variable selection through the use of variable selection networks applied to both static covariates and time-dependent covariates.} \cite{lim_temporal_2020} The variable selection mechanism also allows the model to remove noisy inputs that might have a negative effect on performance.
@@
 
-    \begin{figure}[ht!]
+    \afterpage{%
+    \begin{figure}
         \centering
-        \includegraphics[scale=0.6]{imgs/solar_TFT_overview.jpg}
+        \makebox[\textwidth][c]{\includegraphics[scale=0.6]{imgs/solar_TFT_overview.jpg}}
         \caption{An overview of the setup of the neural network.
@@
     \end{figure}
+    \clearpage
+    }
 
@@
 
+    \pagebreak
     \subsubsection{Tuning}
@@
 
+    \pagebreak
     \subsubsection{Loss function}\label{sec:loss_function}

### 2022-05-13 MSC-GUDMUNDSSON-2022.tex
Finished thesis. Implemented final feedback from instructor.

@@
 %\whensigned{day}{month}{year} %day is only used on some formats, but you must put something.
-\whensigned{10}{\MLapr}{2022}
+\whensigned{10}{\MLmay}{2022}
 
@@
 We base this on Belgingur ehf.'s high-resolution weather forecast data, the 24-hour forecast irradiance values for the area of the power station.
-Our model achieves a very low average mean absolute percentage error of 13.21\% on the test dataset for one-day forecasts.
+Our model achieves excellent results for one-day forecasts on the test dataset, with an average mean absolute percentage error of 13.21\%.
 In addition, the model outputs a confidence level for its predictions providing insight to a grid operator on whether to expect unpredictability in the power output.
@@
 {
-Hraðar og hraðar erum við að færast frá því að framleiða raforku með jarðefnaeldsneyti, í átt að umhverfisvænni mátum til raforkuframleiðslu eins og sólarorku.
+Heimurinn er að færast frá því að framleiða raforku með jarðefnaeldsneyti, í átt að umhverfisvænni leiðum til raforkuframleiðslu eins og sólarorku.
 Sólarorka er óútreiknanleg og mjög háð veðurfari, aðallega skýjahulu.
-Óstöðugleiki og óútreiknanleiki eru óákjósanlegir þegar kemur að sólarorku, þar sem framleiðsla verður að vera jöfn neyslu.
-Til að tryggja afhendingu raforku, þegar sólarorka er notuð, reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði.
-Áreiðanleg geta til að sjá fyrir um sólarorkuframleiðslu gerir stjórnun raforkukerfa einfaldari og útreiknanlegri, sem eykur skilvirkni þeirra.
-Verðmæt gögn eru til í formi veðurspáa, sem hægt er að nýta til að gera sólarorkuframleiðslu útreiknanlegri.
+Óstöðugleiki og óútreiknanleiki eru óæskilegir þegar kemur að sólarorku, þar sem framleiðsla verður að vera jöfn neyslu.
+Til að tryggja afhendingu raforku þegar sólarorka er notuð, reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði.
+Sólarorkuframleiðsla yrði skilvirkari og áreiðanlegri ef hægt væri að spá nákvæmar um framleiðslugetu dagsins, t.d.
+með áreiðanlegri veðurspám og áhrif þeirra á framleiðslugetu.
+Verðmæt gögn eru til í formi veðurspáa sem hægt er að nýta til að gera sólarorkuframleiðslu útreiknanlegri.
 Við notum nýjar aðferðir á sviði djúpnáms, athygli(e.
 attention), og þá sérstaklega tímasamrunabylta (e.
-temporal fusion transformers) til að spá fyrir um sólargeislun við sólarorkuver, sem fylgir náið raforkuframleiðslu orkuversins.
+temporal fusion transformers) til að spá fyrir um sólargeislun við sólarorkuver sem fylgir náið raforkuframleiðslu orkuversins.
 Við byggjum þessa vinnu á háupplausnarveðurspám Belgings ehf., 24 klst.
@@
 RNN &Recurrent Neural Network\\
+CNN &Convolutional Neural Network\\
 GRU &Gated Recurrent Unit\\
@@
 TCNN &Temporal Convolutional Neural Network\\
-NARXNN &Non-linear Auto Regressive Neural Network\\
+NARXNN &Non-linear Auto-Regressive Neural Network\\
 TFT &Temporal Fusion Transformer\\

### 2022-05-13 background.tex
Finished thesis. Implemented final feedback from instructor.

@@
 \chapter{Background\label{cha:background}}
-In this chapter, we discuss the work our work builds on to give our work context.
+In this chapter, we discuss the work we build on.
 We cover the history of neural networks in machine learning, previous architectures that have been used to solve similar problems and the components of the solution we arrived at.
+We also give a summary of weather forecasting and its use in both our work and previous methods for solar power prediction.
 
@@
 Artificial Neural Networks (ANNs), generally known as just Neural Networks (NNs) were first theorized in the 1940s \cite{mcculloch_logical_1943}.
-They are supposed to emulate the function of biological brains.
+They are supposed to loosely emulate the function of biological brains.
 They are composed of \enquote{neurons} which, when they receive a signal, apply a function to the signal which they then relay to the next neuron.
@@
 These values are then passed on to the next layer.
-Training a neural network is the process of optimizing the weights in the layers of the network such that a specified input yields the desired output via iteration to minimize the output of a loss function \cite{goodfellow_deep_2016}.\\
+Training a neural network is the process of optimizing the weights in the layers of the network such that a specified input yields the desired output via iteration to minimize the output of a loss function \cite{goodfellow_deep_2016}.
+
 Frank Rosenblatt created the perceptron in the 1950s.
@@
 His training algorithm was very limited and so inefficient that it did not allow for neural networks with more than one layer \cite{rosenblatt_perceptron_1958}.
-Based on his work, Rosenblatt first theorized the ability of neural networks to recognize people and translate between languages \cite{vandeput_brief_nodate}.\\
+Based on his work, Rosenblatt first theorized the ability of neural networks to recognize people and translate between languages \cite{vandeput_brief_nodate}.
+
 In 1974 Paul Werbos proposed backpropagation to optimize neural networks \cite{Werbos:74}, which became popular in 1986  with the work of Rumelhart et al.
@@
 \enquote{The backpropagation algorithm works by computing the gradient of the loss function with respect to each weight by the chain rule, computing the gradient one layer at a time, iterating backward from the last layer to avoid redundant calculations of intermediate terms in the chain rule} \cite{noauthor_backpropagation_2022}.
-Backpropagation enabled the creation of deeper neural networks, notably used to read handwritten digits with good accuracy \cite{lecun_backpropagation_1989}.\\
+Backpropagation enabled the creation of deeper neural networks, notably used to read handwritten digits with good accuracy \cite{lecun_backpropagation_1989}.
+
 Although backpropagation enabled neural networks with multiple layers, it did not enable significant depth.
@@
 This slows down the gradient descent used to optimize the networks, also at an exponential rate.
-In the 2010s the ReLu activation function \cite{nair_rectified_2010} coupled with improved optimization algorithms like ADAM and massive advances in computer hardware opened the door for Deep Learning, neural networks with many more layers.
-Today, deep neural networks have been used to great effect in applications like Google DeepMind's AlphaGO \cite{silver_mastering_2016}, the first AI to beat the world champion in the game of go, and OpenAI's GPT3 model generates text human-like enough to pass the Turing test, to be confused with text written by a human \cite{vandeput_brief_nodate}.
+In the 2010s the ReLu activation function \cite{nair_rectified_2010} coupled with improved optimization algorithms like ADAM \cite{kingma_adam_2014} and massive advances in computer hardware opened the door for Deep Learning, neural networks with many more layers.
+Today, deep neural networks have been used to great effect in applications like Google DeepMind's AlphaGO \cite{silver_mastering_2016}, the first AI to beat the world champion in the game of Go, and OpenAI's GPT3 model generates text human-like enough to pass the Turing test, to be confused with text written by a human \cite{vandeput_brief_nodate}.
 
 \section{Recurrent Neural Networks\label{cha:RNN}}
-Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) and Long Short-Term Memory (LSTM) cells \cite{Goodfellow-et-al-2016, hochreiter_long_1997} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
+Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) \cite{cho_properties_2014} and Long Short-Term Memory (LSTM) cells \cite{hochreiter_long_1997, Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
 RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently \cite{lim_time_2021}.
 LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.
-Figure~\ref{fig:RNNs} shows the structure of the nodes of the three different Recurrent Neural Network types discussed above.\\
+Figure~\ref{fig:RNNs} shows the structure of the nodes of the three different Recurrent Neural Network types discussed above.
 
@@
 $\sigma$ and $tanh$ are activation functions.
-\cite{lopez_rnn_nodate}.
+\cite{lopez_rnn_nodate}
     \label{fig:RNNs}}
@@
     \includegraphics[scale=0.80]{imgs/transformer.png}
-    \caption{The transformer architecture.\cite{vaswani_attention_2017}
+    \caption{The transformer architecture. \cite{vaswani_attention_2017}
     \label{fig:transformer}}
@@
 
-Transformers have been used for a while in Natural Language Processing (NLP) and have more recently been used very effectively in forecasting \cite{vaswani_attention_2017}.
+Transformers have been used for some time in Natural Language Processing (NLP) and have more recently been used very effectively in forecasting \cite{vaswani_attention_2017}.
 They can more effectively learn and preserve local context located far apart in the input data.
@@
     \includegraphics[scale=0.5]{imgs/encoder_decoder.jpeg}
-    \caption{The encoder/decoder structure \cite{nechu_what_2020}
+    \caption{The encoder/decoder structure. \cite{nechu_what_2020}
     \label{fig:encoder_decoder}}
@@
 \subsection{Positional encoding}
-Unlike RNNs, the transformer architecture has no recurrence, or cell based memory, and does not inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
+Unlike RNNs, the transformer architecture has no recurrence, or cell-based memory, and does not inherently track the order of the input data, so positional information has to be injected into the network input for the model to be able to use the positional context of the data.\cite{vaswani_attention_2017}
 
@@
 Numerical Weather Prediction (NWP) has used mathematical models of the oceans and atmosphere since the 1920s to predict the weather based on current conditions, although realistic results weren't achieved until the advent of computer simulations in the 1950s \cite{noauthor_numerical_2022}.
-The Weather Research and Forecasting (WRF) model is an NWP system created by the National Center for Atmospheric Research (NCAR), the National Oceanic and Atmospheric Administration (NOAA), the Air Force Weather Agency (AFWA), the Naval Research Laboratory (NRL), the University of Oklahoma (OU), and the Federal Aviation Administration (FAA) in the United States for both research and operational weather forecasting \cite{noauthor_weather_2022}.\\
-WRF-SOLAR is an NWP model based on WRF specifically designed to model values useful for solar power, including high frequency irradiance calculations, more accurate solar position algorithms, and more robust aerosol and particle simulations with regard to radiation \cite{jimenez_wrf-solar_2016}.
+The Weather Research and Forecasting (WRF) model is an NWP system created by the National Center for Atmospheric Research (NCAR), the National Oceanic and Atmospheric Administration (NOAA), the Air Force Weather Agency (AFWA), the Naval Research Laboratory (NRL), the University of Oklahoma (OU), and the Federal Aviation Administration (FAA) in the United States for both research and operational weather forecasting \cite{noauthor_weather_2022}.
+
+WRF-SOLAR is an NWP model based on WRF specifically designed to model values useful for solar power, including high-frequency irradiance calculations, more accurate solar position algorithms, and more robust aerosol and particle simulations with regard to radiation \cite{jimenez_wrf-solar_2016}.
 \section{Previous work}
 Lin et al.
-\cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on solar power prediction, Temporal Convolutional Neural Networks (TCNN) can show even better performance.
+\cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on solar power prediction, Temporal Convolutional Neural Networks (TCNN) can achieve even better performance.
 \textit{\endquote{TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
-and residual connections}} \cite{lin_temporal_2020}. \\
+and residual connections}} \cite{lin_temporal_2020}.
+
 Jaidee et al.
@@
 They generate multiple candidate neural networks and use the best networks as a base to generate new networks for the next generation.
-This is done for multiple generations and in the end, the best neural network generated is used \cite{jaidee_very_2019}.\\
+This is done for multiple generations and in the end, the best neural network generated is used \cite{jaidee_very_2019}.
+
 Su et al.
-\cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto Regressive Neural Networks (NARXNN), as well as various statistical approaches.
-When predicting solar power output, NARXNN takes in the current power output values and considers them, in real time, as well as the past values of the power output of the system.
+\cite{su_machine_2019} did extensive experimentation on more conventional neural networks and newer, less known networks including  Non-linear Auto-Regressive Neural Networks (NARXNN), as well as various statistical approaches.
+When predicting solar power output, NARXNN takes in the current power output values and considers them, in real-time, as well as the past values of the power output of the system.
 These experiments showed that NARXNN had the best performance of these, by a rather large margin, and a hybrid method, of the better performing methods, showed even better results \cite{anderson_using_2018}.
 The statistical methods generally performed significantly worse in predicting the power output.
-\\
+
 Satellite imagery has been used for predicting solar power output as well.
-A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.\\
+A neural network was trained to predict solar power production, primarily from cloud cover information it learned from these satellite images \cite{jang_solar_2016}.

### 2022-05-13 discussion.tex
Small fixes
Finished thesis. Implemented final feedback from instructor.

@@
 The downward diffuse shortwave irradiance, the other network input data value, is a much smaller part of the total irradiance and has, to the bare eye, seemingly no correlation with the observation values.
-Despite this, the models show generally fantastic performance in predicting the irradiance.
+Despite this, the models show generally excellent performance in predicting the irradiance.
 The full model was able to utilize the forecast data from the surrounding area to achieve noticeably better performance.
@@
 We do not know how close this model comes to maximizing the information available in the input data.
+
+\pagebreak
 
@@
 
-\section{Limitations and Future Work}
+\section{Limitations and future work}
 We used a dataset of a fairly limited size, only 425 days.
@@
 
-Sadly, there is not a standard benchmark for solar irradiance forecast models, and the data used by others who have done similar work is not available, so we can't make a proper comparison between methodologies and models.
+Sadly, there is not a standard benchmark for solar irradiance forecast models, and the data used by others who have done similar work is not available, so we can not make a proper comparison between methodologies and models, as weather predictability can vary hugely between locations.
+However, for some reference, Lee et al.
+produced a combination Convolutional Neural Network (CNN) and Long Short-Term Memory (CNN) model which managed to forecast the power output of a solar power station in Scotland 6 hours into the future with a MAPE of 25.85\% \cite{lee_forecasting_2018}.
 
 \section{Conclusion\label{sec:conclusions}}
-The results clearly show that relying on forecast irradiance data from meteorological models is a very effective, but not perfect, method of accurately predicting irradiance in a single area.
-A transformer model, the Temporal Fusion Transformer, managed to very effectively translate patterns in meteorological irradiance forecasts to accurate local irradiance predictions utilizing spatially and temporally adjacent data.\\
+The results clearly show that relying on forecast irradiance data from meteorological models is an effective, but not perfect, method of accurately predicting irradiance in a single area.
+A transformer model, the Temporal Fusion Transformer, managed to effectively translate patterns in meteorological irradiance forecasts to accurate local irradiance predictions utilizing spatially and temporally adjacent data.\\
 %%% Local Variables: 

### 2022-05-13 introduction.tex
Small fixes
Finished thesis. Implemented final feedback from instructor.

@@
 Fossil fueled power stations are generally kept on standby, waiting to fulfil this demand \cite{lee_forecasting_2018}.
-These power stations often take hours to start, generally pollute more than other power stations and are expensive to run \cite{lee_forecasting_2018}.\\
+These power stations often take hours to start, generally pollute more than other power stations, and are expensive to run \cite{lee_forecasting_2018}.\\
 Reliable and accurate forecasts of the fluctuations in solar power production can give power grid operators notice and insight into future power production \cite{lee_forecasting_2018}.
@@
 
-Here we explore using high-resolution weather forecast data from meteorological models, offloading a majority of this difficult task to the field of meteorology.
+Here we explore using high-resolution weather forecast data from meteorological models of irradiance, to improve the predictions  of solar power production.
+More specifically offloading a majority of this difficult task to the field of meteorology.
 Instead of generating forecasts, we focus on correlating existing forecasts with measurements and detecting and reducing temporal and spatial error in the general forecasts using newer methods like Temporal Fusion Transformers.
-What is the quality of the predictions this methodology produces? Can we create a model which is useful to power producers in operating their systems more efficiently?\\
+This thesis aims to address two questions: What is the quality of the predictions this methodology produces?
+Can we create a model which is useful to power producers in operating their systems more efficiently?\\
 
 This thesis is structured as follows.
-First, we talked about the motivations for the work.
-Next, we discuss the background the work stands on.
+First, we talk about the motivations for the work.
+Next, we discuss the background that the work stands on.
 After that, we cover the methods utilized, the data and tools used to achieve these goals.
-The proceeding chapter contains the results of the work, and finally, we discuss the implications, strengths and flaws of the work, future work and end with conclusions.
+The following chapter contains the results of the work, and then we discuss the implications, strengths and weaknesses of the work.
+Finally, we conclude and discuss future work.
 

### 2022-05-13 methods.tex
Finished thesis. Implemented final feedback from instructor.

@@
 \subsection{Irradiance measurements}
-Under certain circumstances irradiance values simulated after the fact can be accurate, however this does not apply in all conditions, and real measured values are much more reliable \cite{gueymard_global_2008}.
+Under certain circumstances irradiance values simulated after the fact can be accurate, however, this does not apply in all conditions, and real measured values are much more reliable \cite{gueymard_global_2008}.
 The irradiance was measured with seven Kipp \& Zonen CM21 pyranometers placed around the power station.
@@
 
-\subsection{Forecast Data\label{cha:forecast_data}}
+\subsection{Forecast data\label{cha:forecast_data}}
 The weather forecast data is provided by Belgingur ehf., Reykjavík, Iceland.
@@
 We use two irradiance values from the forecast, the downward irradiance and the downward diffuse irradiance.
-These irradiance values were normalized linearly to the range $0.0-1.0$ before being input into the neural network.
+These irradiance values were also normalized linearly to the range $0.0-1.0$ before being input into the neural network.
 
@@
 
-    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data, or datasets where it is unknown which parts are of significance.\\
+    \noindent\textbf{Gating mechanisms} identify and skip unused input data, granting models greater ability to work with less curated input data or datasets where it is unknown which parts are of significance.
+
     The Gated Residual Network (GRN) structure, shown in Figure~\ref{fig:ttgt_detail}, is a base building block of TFTs \cite{lim_temporal_2020}.
-It provides the model the flexibility to choose the complexity of its input, more complex where needed, and less complex where not.\\
+It provides the model with the flexibility to choose the complexity of its input, more complex where needed, and less complex where not.
+
     \clearpage
@@
          \makebox[\textwidth][c]{\includegraphics[scale=0.36]{imgs/TFT.png}}
-        \caption{A detailed view of the architecture of the Temporal Fusion transformer.\cite{lim_temporal_2020}
+        \caption{A detailed view of the architecture of the Temporal Fusion transformer. \cite{lim_temporal_2020}
         \label{fig:ttgt_detail}}
@@
 Since the goal of tuning was not only to minimize Mean Absolute Percentage Error but balance that with the ability to recognize uncertainty, the process was very manual.
-Tables~\ref{tab:full_params} and \ref{tab:lim_params} show the selected hyperparameters for the full and limited models respectively.\\
+Tables~\ref{tab:full_params} and \ref{tab:lim_params} show the selected hyperparameters for the full and limited models respectively.
 
@@
     \subsubsection{Loss function}\label{sec:loss_function}
-    The loss function used for Temporal Fusion Transformers is quantile loss, as it's an integral part of generating the confidence intervals of the model \cite{lim_temporal_2020}.
+    The loss function used for Temporal Fusion Transformers is quantile loss, as it is an integral part of generating the confidence intervals of the model \cite{lim_temporal_2020}.
     A model's prediction is a random value within a probability distribution \cite{wen_multi-horizon_2018}.

### 2022-05-13 results.tex
Small fixes
Finished thesis. Implemented final feedback from instructor.

@@
 We show irradiance plots for select days representative of the models' performance in various conditions.
-These plots show the measured irradiance for the day, the models' predicted irradiance values, along with a 90\% confidence interval, detailed in section~\ref{sec:loss_function}.
+These plots show the measured irradiance for the day, the models' predicted irradiance values, along with a 90\% confidence interval, detailed in Section~\ref{sec:loss_function}.
 The plots also show the forecast downward shortwave irradiance, the largest factor in measured irradiance, from the model input.
@@
 In Figure~\ref{fig:full_med_med} we see examples of days where the model achieves higher error, MAPE in the range 19-35\%, but with generally larger 90\% confidence intervals so the model still keeps the measured values close to or within them.
-\begin{figure}[ht!]
+
+\afterpage{%
+\begin{figure}
     \centering
@@
 Figures~\ref{fig:full_med_bad_a} and \ref{fig:full_med_bad_d} show where the model predicts a constant lower irradiance when half the day had high irradiance and the other half low.
-Figures~\ref{fig:full_med_bad_b} and \ref{fig:full_med_bad_c} show the model predicting a mostly or fully sunny day when it wasn't.
+Figures~\ref{fig:full_med_bad_b} and \ref{fig:full_med_bad_c} show the model predicting a mostly or fully sunny day when it was not.
 In all of these cases, the observed irradiance is far outside the 90\% confidence interval.
-\begin{figure}[ht!]
+
+\afterpage{%
+\begin{figure}
     \centering

### 2022-05-13 MSC-GUDMUNDSSON-2022.tex
Tiny fixes. Thesis in final mode. COMPLETE.

@@
 %% online: for online versions (skip blank pages)
-\documentclass[online,masters,deptscs,forcegraphics,draft]{ruthesis}
+\documentclass[online,masters,deptscs,forcegraphics,final]{ruthesis}
 
@@
 %\whensigned{day}{month}{year} %day is only used on some formats, but you must put something.
-\whensigned{10}{\MLmay}{2022}
+\whensigned{20}{\MLmay}{2022}
 
@@
 \setSemester{2022-2}
-\setShortSignedDate{10.4.2022}
+\setShortSignedDate{20.5.2022}
 
@@
 Óstöðugleiki og óútreiknanleiki eru óæskilegir þegar kemur að sólarorku, þar sem framleiðsla verður að vera jöfn neyslu.
-Til að tryggja afhendingu raforku þegar sólarorka er notuð, reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði.
+Til að tryggja afhendingu raforku þegar sólarorka er notuð reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði.
 Sólarorkuframleiðsla yrði skilvirkari og áreiðanlegri ef hægt væri að spá nákvæmar um framleiðslugetu dagsins, t.d.
@@
 mean absolute percentage error) upp á 13,21\% á prófunargagnasettinu fyrir eins dags spár.
-Að auki skilar módelið öryggisbili fyrir spárnar, sem veitir raforkuframleiðendum innsæi um það hvort þeir geti búist við óútreiknanlegri sólarorkuframleiðslu.
+Að auki skilar módelið öryggisbili fyrir spárnar sem veitir raforkuframleiðendum innsæi um það hvort þeir geti búist við óútreiknanlegri sólarorkuframleiðslu.
 Módelið skilar nákvæmari spám þegar það styðst við veðurspágögn af stærra svæði í nánd orkuversins en þegar það nýtir eingöngu gögn fyrir nærumhverfi orkuversins.

### 2022-05-13 methods.tex
Tiny fixes. Thesis in final mode. COMPLETE.

@@
 
+\newpage
+
 \section{Model}
@@
          \makebox[\textwidth][c]{\includegraphics[scale=0.5]{imgs/TFT_overview.jpg}}
-        \caption{An overview of the Temporal Fusion Transformer.
+        \caption{An overview of the Temporal Fusion Transformer \cite{lim_temporal_2020}.
         \label{fig:tft_overviwew}}
@@
 
-    \clearpage
     \noindent\textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}.
@@
         \centering
-         \makebox[\textwidth][c]{\includegraphics[scale=0.36]{imgs/TFT.png}}
-        \caption{A detailed view of the architecture of the Temporal Fusion transformer. \cite{lim_temporal_2020}
+         \makebox[\textwidth][c]{\includegraphics[scale=0.3]{imgs/TFT.png}}
+        \caption{A detailed view of the architecture of the Temporal Fusion transformer \cite{lim_temporal_2020}.
         \label{fig:ttgt_detail}}
@@
         \centering
-        \makebox[\textwidth][c]{\includegraphics[scale=0.6]{imgs/solar_TFT_overview.jpg}}
+        \makebox[\textwidth][c]{\includegraphics[scale=0.5]{imgs/solar_TFT_overview.jpg}}
         \caption{An overview of the setup of the neural network.

### 2022-05-16 background.tex
Tiny fixes. Thesis in final mode. COMPLETE.
Fixed typo which made quotation marks disappear

@@
 $x_t$ represents inputs, $h_t$ and $C_t$ represent hidden layer vectors passed between cells.
-$\sigma$ and $tanh$ are activation functions.
-\cite{lopez_rnn_nodate}
+$\sigma$ and $tanh$ are activation functions \cite{lopez_rnn_nodate}.
     \label{fig:RNNs}}
@@
     \centering
-    \includegraphics[scale=0.80]{imgs/transformer.png}
-    \caption{The transformer architecture. \cite{vaswani_attention_2017}
+    \includegraphics[scale=0.70]{imgs/transformer.png}
+    \caption{The transformer architecture \cite{vaswani_attention_2017}.
     \label{fig:transformer}}
@@
 
+\newpage
+
 \subsection{Encoder/decoder}
@@
     \includegraphics[scale=0.5]{imgs/encoder_decoder.jpeg}
-    \caption{The encoder/decoder structure. \cite{nechu_what_2020}
+    \caption{The encoder/decoder structure \cite{nechu_what_2020}.
     \label{fig:encoder_decoder}}
@@
 \cite{lin_temporal_2020} showed that while the more traditional GRUs and LSTMs show good performance on solar power prediction, Temporal Convolutional Neural Networks (TCNN) can achieve even better performance.
-\textit{\endquote{TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
+\enquote{\textit{TCNN is a novel convolutional architecture designed for sequential modelling, which combines causal and dilated convolutions
 and residual connections}} \cite{lin_temporal_2020}.

### 2022-05-20 background.tex
Minor fix. Final version for final handin. Defence complete.

@@
 \section{Recurrent Neural Networks\label{cha:RNN}}
-Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) \cite{cho_properties_2014} and Long Short-Term Memory (LSTM) cells \cite{hochreiter_long_1997, Goodfellow-et-al-2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
+Recurrent Neural Networks (RNNs), especially Gated Recurrent Units (GRUs) \cite{cho_properties_2014} and Long Short-Term Memory (LSTM) cells \cite{hochreiter_long_1997, goodfellow_deep_2016} along with more advanced derivatives have been shown to be able to achieve good results when trained on previous weather and power generation data \cite{lin_temporal_2020, lee_forecasting_2018, jaidee_very_2019, su_machine_2019}.
 RNNs are a class of deep neural networks where nodes are connected in a temporal sequence, allowing the network to learn time-based data efficiently \cite{lim_time_2021}.
-LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{Goodfellow-et-al-2016, noauthor_recurrent_2021}.
+LSTMs and GRUs are types of RNNs in which, the nodes in the network have additional stored states controlled by the network via mechanisms that use time delays or feedback loops \cite{goodfellow_deep_2016, noauthor_recurrent_2021}.
 Figure~\ref{fig:RNNs} shows the structure of the nodes of the three different Recurrent Neural Network types discussed above.

### 2022-05-20 results.tex
Tiny fixes. Thesis in final mode. COMPLETE.
Absolute final formatting for final handin

@@
     \centering
-    \includegraphics[scale=1]{imgs/graphs/full/days_full.png}
+    \includegraphics[scale=0.8]{imgs/graphs/full/days_full.png}
     \caption{The recorded Mean Absolute Percentage Error (MAPE) for each day in the test dataset for the full model. 
@@
     \centering
-    \includegraphics[scale=1]{imgs/graphs/less/days_single.png}
+    \includegraphics[scale=0.8]{imgs/graphs/less/days_single.png}
     \caption{The recorded Mean Absolute Percentage Error for each day in the test dataset of the limited model. 
@@
 
-\newpage
 \subsection{Low variance days}
@@
 
-\afterpage{%
-\vspace{5cm}
-\begin{figure}
+\begin{figure}[ht!]
     \centering
-    \makebox[\textwidth][c]{\subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}\label{fig:less_med_a}}\qquad}
-    \makebox[\textwidth][c]{\subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_2.png}\label{fig:less_med_b}}\qquad}
-    \makebox[\textwidth][c]{\subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_1.png}\label{fig:less_med_c}}\qquad
-    }
+    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_g_1.png}\label{fig:less_med_a}}\qquad
+    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_2.png}\label{fig:less_med_b}}\qquad
+    \subfloat[]{\includegraphics[scale=0.5]{imgs/graphs/less/medium/med_m_1.png}\label{fig:less_med_c}}\qquad
     \caption{Examples of days with some drops in irradiance.
@@
 \end{figure}
-\clearpage
-}
 
